### 2021-08-09 docs/09-09-21_three-category/three-category.Rmd
Slides up

---
title: A multinomial spatio-temporal model for sexual risk behaviour in AGYW
subtitle: HIV Inference Lab Group Meeting
author: Adam Howes
institute: Imperial College London
date: August 2021
bibliography: citations.bib
output:
  beamer_presentation:
    latex_engine: pdflatex
    highlight: haddock
    fig_width: 7 
    fig_height: 3
    includes:
        in_header: preamble.tex
---

## Background

* Want to estimate the proportion $p_k$ of AGYW in $K = 4$ risk categories
  * $k = 1$: Not sexually active `nosex12m`
  * $k = 2$: Cohabiting, single partner `sexcohab`
  * $k = 3$: Non-regular partner(s) `sexnonreg`
  * $k = 4$: Key population, female sex worker `sexpaid12m`
* Used as a step towards estimating total incidence by group
  * By district, age, sex, time
* Different policy responses depending on where new infections are

## Goal

* Estimate proportions indexed by
  * District $i = 1, \ldots, n$
    * 13 AGYW Global Fund priority countries (but fitting individually): Botswana, Cameroon, Kenya, Lesotho, Malawi, Mozambique, Namibia, South Africa, Swaziland, Tanzania, Uganda, Zambia and Zimbabwe
  * Survey $t = 1, \ldots, T$
    * The available DHSs
  * Age $a \in \{\text{15--19}, \text{20--24}, \text{25--29}\}$
  * Category $k$
* e.g. in Lilongwe, in 2015, how many AGYW aged 20-24 are cohabiting with a single partner?

## Why not just use the raw data

1. Usual high variance from low counts problem, improved by spatio-temporal smoothing
2. The raw proportions don't add up to one
  * This is illustrated in the next slide

##

:::: columns

::: column

```{r echo=FALSE}
knitr::include_graphics("page1-check100.pdf")
```

:::

::: column

```{r echo=FALSE}
knitr::include_graphics("page2-check100.pdf")
```

:::

::::

## Multinomial likelihood

* We have proportions $p_1, \ldots, p_4$ which are all $\in [0, 1]$ and $\sum_k p_k = 1$
* $\implies$ suggests using a multinomial likelihood $\mathbf{y}_{ita} \sim \text{Multinomial}(m_{ita}; \, p_{ita1}, \ldots, p_{itaK})$
$$
p(\mathbf{y}_{ita} \, | \, m_{ita}, p_{ita1}, \ldots, p_{ita4}) = \frac{m_{ita}!}{y_{ita1}! \times \cdots \times y_{ita4}!} p_{ita1}^{y_{ita1}} \times \cdots p_{ita4}^{y_{ita4}}
$$

## Multivariate model

* Requires a multivariate model in that each observation depends on $K$ probabilities
  * Really $K - 1$ numbers, since they are constrained to sum to one
* Usually this is done by modelling the "contrasts" between categories
  * How much more likely to be in this category (or group of categories) than that category (or group of categories)
  * Different ways to do this
* These types of models are popular in economics to model choice
  * Treat the latent field like utility
* Some keywords: discrete choice model, multinomial logit model

## Baseline category

* Pick one of the $k \in \{1, \ldots, K\}$ as your baseline category, say $k = 2$
* Define log-odds for being in category $k \neq 2$ versus $k = 2$
$$
\log \left( \frac{p_{itak}}{p_{ita2}} \right)
$$
* Three of these: $\{2\}$ versus $\{1\}$, $\{2\}$ versus $\{3\}$ and $\{2\}$ versus $\{4\}$

## Nested

* Attractive when it is appropriate to model individuals as making choices sequentially
* $\{1\}$ versus $\{2, 3, 4\}$, $\{2\}$ versus $\{3, 4\}$ and $\{3\}$ versus $\{4\}$
* Choosing 1. whether or not to have sex, 2. conditional on having sex, whether to have one cohabiting partner or irregular partners seems reasonable
* Choosing 3. whether or not to be a FSW based upon choosing to have irregular partners is less reasonable
  * i.e. not like you first choose to have many partners then choose to be FSW, more like choose to be FSW based on economic factors say

## Workaround requried

* Want to fit this model quickly and easily using `R-INLA`
* Sadly `R-INLA` doesn't work for likelihoods which depend upon multiple elements of the latent field (like this multinomial model)
* There is a workaround, called the Poisson trick
  * A multinomial likelihood can be rewritten as a Poisson likelihood (with some additional nuisance parameters)
  * So this respecifies the model in terms of a Poisson likelihood for each category of the observation

## Sidenote: ordinal

* $\pi_{k} = \sum_{l \leq k} p_{l}$ be the cumulative probability of category $k$ where $\pi_{1} \leq \pi_{2} \leq \cdots \leq \pi_{K}$.
* Ordinal logistic regression is based upon $K - 1$ cumulative logits $\text{logit} (\pi_{k}) = \eta_{k}$ where the linear predictors are ordered
* One way to do this is to assume $\eta_k$ only differ in their intercept (also called cut-point)
* I wonder if you can fit this with a single linear predictor and a category random effect?
  * Possibly wouldn't need Poisson trick

## Poisson trick

* Based on the following fact
* Let $y_k \sim \text{Pois}(\lambda_k)$, $k = 1, \ldots, K$ then $\y \, | \, n \sim \text{Multinomial}(\y; n, p_1, \ldots, p_K)$
  * $m = \sum_k y_k$
  * $p_j = \lambda_j / \sum_{k = 1}^K \lambda_k$
* In words, given their sum, Poisson counts are jointly multinomially distributed [@mccullagh1989generalized]
  * In some some way you can obtain the multinomial likelihood (what we want) from Poisson likelihoods (what we have to work with)

## Poisson trick

* In the multinomial likelihood, the sample size is treated as fixed
* Instead treat $m = \sum_k y_k$ as random. Let $\Lambda = \sum_{k = 1}^K \lambda_k$ and suppose that $m \sim \text{Poisson}(\delta \Lambda)$ [@lee2017poisson]
\begin{align*}
P(\mathbf{y}, m) = P(m)P(\mathbf{y} \, | \, m) &= \exp(-\delta \Lambda) \frac{(\delta \Lambda)^m}{m!} \times \frac{m!}{\prod_k y_k!} \prod_k \left( \frac{\lambda_k}{\Lambda} \right)^{y_k} \\
&= \prod_k \left( \frac{\exp(-\delta \lambda_k) \left( \delta \lambda_k \right)^{y_k}}{y_k!} \right)
\end{align*}

## Poisson trick

* To find the marginal $P(\mathbf{y})$ sum over the possible values of $m$ (0 to $\infty$)
$$
P(\mathbf{y}) = \prod_k \left( \frac{\exp(-\delta \lambda_k) \left( \delta \lambda_k \right)^{y_k}}{y_k!} \right)
$$
* $\implies y_k \sim \text{Poisson}(\delta \lambda_k)$ independently

## Poisson trick

* "The Poisson surrogate model eliminates $m$ from the denominator of the multinomial probabilities. This makes sense intuitively, as we do not expect the multinomial sums to provide any useful information in estimating the fixed effects." [@lee2017poisson]
* Example of a broader class of Poisson transformations where the theme is to consider the normalisation constant as another parameter [@barthelme2015poisson]
* I guess that this must mean the normalisation constant is "estimated" exactly always

## Poisson log-linear model

* $y_{itak} \sim \text{Poisson}(\lambda_{itak})$
* $\log(\lambda_{itak}) = \eta_{itak}$ the linear predictor for district $i$, survey $t$, age-group $a$ and category $k$
* Including observation-specific random effects $\theta_{ita} \sim \mathcal{N}(0, \tau_\theta^{-1})$ in the model $\eta_{itak} = \theta_{ita} + \cdots$ assures exact reproduction of the multinomial denominators
  * By "observation" I mean observation of the multinomial e.g. a vector `c(2, 5, 3, 0)`
  * These variables correspond to $\log(\delta)$ from before. So we're just having a free parameter which can 
* Then you put whatever mixed model things you'd like in

## No intercept

* Don't include an intercept, i.e. `-1` in the formula
* This is because we are interested in recovering probabilities from the model
$$
p_j = \frac{\lambda_j}{\sum_{k = 1}^K \lambda_k} = \frac{\exp{\eta_j}}{\sum_{k = 1}^K \exp{\eta_j}}
$$
* So adding a constant $\beta_0$ on to each of the $\eta_j$ does nothing (so I'd hope that if you did include a constant it would turn out to be zero if you fitted the model)
$$
\frac{\exp{(\eta_j + \beta_0)}}{\sum_{k = 1}^K \exp{(\eta_j + \beta_0)}} = \frac{\exp{\eta_j}}{\sum_{k = 1}^K \exp{\eta_j}}
$$
* Note that this `exp(x) / sum(exp(x))` function is called "softmax"

## Nine models

* Always use age-category random effects $\alpha_{ak} \sim \mathcal{N}(0, \tau_\alpha^{-1})$
  * I haven't been including category random effects, but based on my research for this presentation I think I should have been. I think because I have the age-category it's already mopping up the intercept
* Three choices for spatial-category random effects $\phi_{ik}$:
  * None
  * IID $\phi_{ik} \sim \mathcal{N}(0, \tau_\phi^{-1})$
  * BYM2
* Three choices of time-category random effects $\gamma_{tk}$
  * None
  * IID $\gamma_{tk} \sim \mathcal{N}(0, \tau_\gamma^{-1})$
  * AR1
* Some amount of nuance / choice to defining the structured {spatial (BYM2), temporal (AR1)}-category interactions

## First way

* Let `area_idx` be an indicator the the area
* `area_idx.k = ifelse(category == k, area_idx, NA)` is that same indicator but just for the category `k`
* Then (say you want spatial) you can use `f(area_idx.k, model = "bym2", graph = adjM)` in your `R-INLA` formula
* This results in a set of $K$ independent spatial random effects
$$
\phi_{ik} = \sqrt{\tau_{\phi_{k}}^{-1}} \left(\sqrt{\pi_k} \cdot u_{ik} + \sqrt{1 - \pi_k} \cdot v_{ik} \right), \quad k = 1, \ldots, K
$$
* Key point is that you have different hyperparameters for each category $\{\tau_{\phi_k}, \pi_k\}$
  * Do we want this?
* Also have to define `area_idx.k` and write `f(...)` more times (annoying)

## Second way

* Alternative (in `R-INLA`) is to use the `group` option which lets you define "Gaussian Kronecker product Markov random fields"
  * Sounds fancy, isn't that fancy
* "Precison matrix" (probably it doesn't have full rank so it's not really a precision) defined by $\mathbf{Q} = \mathbf{Q}_1 \otimes \mathbf{Q}_2$
* In `R-INLA` it's more like the "structure matrix" is defined by $\mathbf{R} = \mathbf{R}_1 \otimes \mathbf{R}_2$
  * So that the precision is $\mathbf{Q} = \tau \mathbf{R}$

## Example of `group` option

* Let's say we're trying to define the time-category interaction random effects, using an AR1 on time
* If there are three surveys, and we suppose $\rho = 1$ then the precision matrix for this AR1 is
$$
\begin{pmatrix}
1 & -1 & 0 \\
-1 & 2 & -1 \\
0 & -1 & 1
\end{pmatrix}
$$
* To get the structure matrix $\mathbf{R_1}$ we'd divide by `exp(mean(log(diag(Q))))`

## Example of `group` option

* For the "category random effects" grouping variable we want IID structure, so that the precision (and structure matrix $\mathbf{R_2}$) are
$$
\begin{pmatrix}
1 & 0 & 0 \\
0 & 1 & 0 \\
0 & 0 & 1
\end{pmatrix}
$$
* The `R-INLA` call for this is `f(sur_idx, model = "ar1", group = cat_idx, control.group = list(model = "iid"), ...)`
* The new structure matrix looks like
$$
\mathbf{R} =
\begin{pmatrix}
\mathbf{R_1} & \mathbf{0} & \mathbf{0} \\
\mathbf{0} & \mathbf{R_1} & \mathbf{0} \\
\mathbf{0} & \mathbf{0} & \mathbf{R_1}
\end{pmatrix}
$$
* Important difference: just have one set of hyperparameters $\{\tau_\gamma, \rho\}$
  * i.e. there is pooling across categories doing it this way

## Survey weights

* Use survey weights to calculate effective observation `x_eff`
* May not be integers i.e. `c(1.87, 5.32, 2.20, 0.15)`
* Deal with this by using `xPoisson` (generalisation of the Poisson to non-integer data) rather than `Poisson`
* Might be equivalent to weighted log-likelihood approach

##

:::: columns

::: column

```{r echo=FALSE}
knitr::include_graphics("page-multinomial-smoothed-district-sexbehav.pdf")
```

:::

::: column

* A (difficult to see) example of what this looks like
* Model 9 has BYM2 and AR1
* Other notes
  * Quantiles produced by manipulating `inla.posterior.sample`s
  * Same samples used in combination with Naomi population data to produce aggregates (i.e. 15-24 age category)

:::

::::

## Data issues for `sexpaid12m` category

* Two different questions asked in the DHS, and answered very differently
* Before 2015, asked about most recent three partners. One of the categories for these is paid
  * Very rare to list one of these as paid, using this question often think there are essentially no FSW
* After 2015, asked if have given or received gifts or money in exchange for sex
  * Better answered but still has issues
    * Missing some FSW from survey sampling frame
    * Not all "gifts of money in exchange for sex" is sex work

##

```{r echo=FALSE}
knitr::include_graphics("../../archive/check_sexpaid-survey-question/20210720-120340-77026581/sexpaid-survey-question.pdf")
```

## Current approach

* Can't fit a (meaningful) spatio-temporal model when the question asked / response is so different
* Have been fitting single-survey spatial models to the most recent DHS
* Otherwise have created a three-category model, moving all `sexpaid12m` into `sexnonreg` to create a new `sexnonregplus` category
  * Haven't checked that some of the `sexpaid12m` shouldn't go into `sexcohab` rather than `sexnonreg` yet (sorry Katie!)
* Challenge is what to do about the `sexpaid12m` category

## What to do about the `sexpaid12m` category

1. Ignore it
2. Try to use other data sources to help
  * Previous estimates from workbook are based upon national estimates of FSW population size (perhaps the UNAIDS Key Population Atlas)

## Johnston et al. (2021, preprint)

**Deriving and interpreting population size estimates for adolescent and young key populations at higher risk of HIV transmission: men who have sex with men, female sex workers and transgender persons**

* Disaggregates the UNAIDS published population size estimates by age (nice, but not by district, which we'd also want) using proportion of sexually active adults
* Kinh is a coauthor and warns that
  * The estimates should (often) be seen as expert opinion rather than based on data
  * Several countries had no data
  * Rounding up when the number is too small

## Laga et al. (2021, preprint)

**Mapping female sex worker prevalence (aged 15-49 years) in sub-Saharan Africa**

* Collates available FSW population size estimates
* Small-area model with covariates used to extrapolate to district level
* Has code available
* Probably the best single source to rely on, done a lot of the work already

## Hodgkins et al. (2021, preprint)

**HIV prevalence, population sizes, and HIV prevention among men who paid for sex in sub-Saharan Africa: a meta-analysis of 82 population-based surveys (2000-2020)**

* Proportion of men who pay for sex (CFSW) estimated from survey data
* Could be linked to proportion FSW by some model e.g.
$$
p_{\text{CFSW}} = B \times p_{\text{FSW}}
$$
with a strongly informative prior $B \sim p(B)$ (around 10 say)
  * Don't know if this is a good model, anyone have any ideas?
  * Extract estimates from Hodgkins and plot versus estimates of FSW to see if it looks linear
  * Of course the ratio is going to vary, but perhaps it's quite stable?
* Fit a model on the good DHS `sexpaid12m` question, then use it where there isn't the question

##

\begin{align}
\log(p_{\text{CFSW}}) - \log(p_{\text{FSW}}) &= \log(B) \\
\log \left( \frac{p_{\text{CFSW}}}{p_{\text{FSW}}} \right) &= \log(B)
\end{align}

## References {.allowframebreaks}