### 2021-09-22 docs/24-09-21_new-cohort/new-cohort.Rmd
Start presentation for new cohort

---
title: "Spatio-temporal statistics for HIV surveillance"
author: "Adam Howes"
date: "August 2021"
output:
  beamer_presentation:
    latex_engine: pdflatex
    highlight: haddock
    fig_width: 7
    fig_height: 3
    includes:
      in_header: preamble.tex
  slidy_presentation:
    highlight: haddock
subtitle: StatML Kick-off Camp
bibliography: citations.bib
institute: Imperial College London
---

## Welcome!

* Welcome to the CDT!
* Congratulations for many of you on completing your previous studies under some non-ideal circumstances

## About me / my project

* Surveillance of HIV in sub-Saharan Africa in collaboration with UNAIDS
* Main interest is in methods and applications of Bayesian statistics towards policy-relevant questions

## Spatio-temporal

* Observations are indexed by space $s$ and time $t$
* For example, often we have a sequence of national surveys
* In social science it's natural that we observe people in space and time
* Key property of spatio-temporal data is correlation structure

## Bayesian statistics

* Bayesian statistics makes sense to me
* Perhaps more importantly, there isn't a lot to remember
  1. Write down a model for your process
  2. Some kind of algorithm to obtain the posterior
  3. Use the posterior for all decision theory

## Data and theory

* The data we work with is incomplete, biased, noisy
* Mainly concerned with getting roughly the right answer
* Happy to use informative priors or scientific knowledge of domain experts to help

## 1. Size of the sample

* Running a (household) survey is expensive
* We only get small-sample sizes, and a low coverage of space-time
* The field of "small-area" estimation should really be called "small-data" estimation
* We use the fact that spatio-temporal data is correlated to improve things with smoothing

## 2. Who is in the sample?

* The sample probably often isn't representative 
* Some people aren't on the sampling frame at all
* Example: for HIV, there are some key population groups (men who have sex with men, transgender people, injecting drug users, sex workers) who are hard to reach with surveys

## 3. Are the answers they give accurate?

* Survey questions can be ambiguous
* People may not be able to or want to answer questions
* Example: age at first sex is an important epidemiological parameter, but men and women (and old and young) give answers which are difficult to make line up

## Sexual risk categories model

* Some work in progress of mine, hopefully will make the above more concrete
* Young women are an important population for HIV prevention

## References {.allowframebreaks}