### 2021-09-24 docs/24-09-21_new-cohort/new-cohort.Rmd
Finish presentation for new cohort

---
title: "Statistics for HIV surveillance"
author: "Adam Howes"
date: "August 2021"
output:
  beamer_presentation:
    latex_engine: pdflatex
    highlight: haddock
    fig_width: 7
    fig_height: 3
    includes:
      in_header: preamble.tex
  slidy_presentation:
    highlight: haddock
subtitle: StatML Kick-off Camp
bibliography: citations.bib
institute: Imperial College London
---

## Welcome!

* Welcome to the CDT!
* Congratulations for many of you on completing your previous studies under some non-ideal circumstances
* PhDs can be challenging, learning process
    * I hope I'm a better researcher than I was two years ago
* Important to find the right work environment, draw on each other

## About me / my project

* Surveillance of HIV in sub-Saharan Africa in collaboration with UNAIDS
* Main interest is in methods and applications of Bayesian statistics towards policy-relevant questions

## Spatio-temporal

* Observations are indexed by space $s$ and time $t$
* For example, often we have a sequence of national surveys
* In social science it's natural that we observe people in space and time
* Key property of spatio-temporal data is correlation structure

## Bayesian statistics

* Bayesian statistics makes sense to me
* Perhaps more importantly, there isn't a lot to remember
  1. Write down a model for your process
  2. Some kind of algorithm to obtain the posterior
  3. Use the posterior for all decision theory

## Data and theory

* The data we work with is incomplete, biased, noisy
* Mainly concerned with getting roughly the right answer
* Happy to use informative priors or scientific knowledge of domain experts to help

## 1. Size of the sample

* Running a (household) survey is expensive
* We only get small-sample sizes, and a low coverage of space-time
* The field of "small-area" estimation should really be called "small-data" estimation
* We use the fact that spatio-temporal data is correlated to improve things with smoothing

## 2. Who is in the sample?

* The sample probably often isn't representative 
* Some people aren't on the sampling frame at all
* Example: for HIV, there are some key population groups (men who have sex with men, transgender people, injecting drug users, sex workers) who are hard to reach with surveys

## 3. Are the answers they give accurate?

* Survey questions can be ambiguous
* People may not be able to or want to answer questions
* Example: age at first sex is an important epidemiological parameter, but men and women (and old and young) give answers which are difficult to make line up

## Sexual risk categories model

* Some work in progress of mine, to hopefully make the above concrete!
* Young women are an important population for HIV prevention
    * 10% of the population and 25% of all new infections [@unaids2021update]
* Not enough resources to provide programs for all young women in high prevalence settings $\implies$ standardised method for accessing population at risk required

##

\begin{table}
\centering
\begin{tabularx}{\textwidth}{lll}
\toprule
Category & HIV related risk & HIV risk ratio \\ 
 \midrule
None & Not sexually active & 0 \\
Low & Sexually active, one cohabiting partner & 1 \\
High & Non-regular sexual partner(s) & 1.72 \\
Very high & Young women from key populations & 13 \\
\bottomrule
\end{tabularx}
\caption{HIV risk categories and risk ratios.}
\end{table}

## Goal

* Estimate proportion of young women in each risk category
  * Districts $i = 1, \ldots, n$ in 13 AGYW Global Fund priority countries
  * Survey $t = 1, \ldots, T$
  * Age $a \in \{\text{15--19}, \text{20--24}, \text{25--29}\}$
* e.g. in Lilongwe (Malawi), in 2015, how many AGYW aged 20-24 are cohabiting with a single partner?

## Spatial random effects

* Tried none, IID and the Besag model (most common structured spatial random effect model)

$$
b_i \, | \, \mathbf{b}_{-i} \sim \mathcal{N} \left(\frac{\sum_{j: j \sim i} b_j}{n_{\delta i}}, \frac{1}{n_{\delta i}}\right),
$$

## Temporal random effects

* Tried none, IID and an AR1 model (most common structured spatial random effect model)

\begin{align*}
a_1 &\sim \left( 0, \frac{1}{1 - \rho^2} \right), \\
a_t &= \rho a_{t - 1} + \epsilon_t, \quad t = 2, \ldots, T, 
\end{align*}

##

\begin{table}[htb]
\centering
\begin{tabularx}{\textwidth}{llllllllll}
Model ID & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 \\ 
\midrule
Spatial structure & S1 & S2 & S3 & S1 & S2 & S3 & S1 & S2 & S3 \\
\midrule
Temporal structure & T1 & T1 & T1 & T2 & T2 & T2 & T3 & T3 & T3 \\
\end{tabularx}
\caption{All of the models considered}
\label{tab:models}
\end{table}

##

```{r echo=FALSE}
knitr::include_graphics("../../archive/process_information-criteria/20210916-133107-9c7e6acd/all-dhs-rank-comparison.pdf")
```

## References {.allowframebreaks}