### 2021-12-31 src/docs_04-01-22-update/04-01-22-update.Rmd
As report is a better way to create slides (e.g. include most recent figures automatically using depends) #50

---
title: A multinomial spatio-temporal model for sexual risk behaviour with application to adolescent girls and young women in 13 sub-Saharan African countries
subtitle: HIV Inference Lab Group Meeting
author: Adam Howes
institute: Imperial College London
date: January 2022
bibliography: citations.bib
output:
  beamer_presentation:
    latex_engine: pdflatex
    highlight: haddock
    fig_width: 7 
    fig_height: 3
    includes:
        in_header: preamble.tex
---

##

```{r echo=FALSE}
knitr::include_graphics("depends/available-surveys.pdf")
```

##

```{r echo=FALSE}
knitr::include_graphics("depends/3-rank-comparison.pdf")
```

##

```{r echo=FALSE}
knitr::include_graphics("depends/model-direct-benefits.pdf")
```

##

```{r echo=FALSE}
knitr::include_graphics("depends/continental-map.pdf")
```

##

```{r echo=FALSE}
knitr::include_graphics("depends/variance-proportions.pdf")
```

##

```{r echo=FALSE}
knitr::include_graphics("depends/within-between-country-variation.pdf")
```

##

```{r echo=FALSE}
knitr::include_graphics("depends/age-variation.pdf")
```

## References {.allowframebreaks}