### 2022-01-27 src/docs_03-02-22-ea-talk/03-02-22-ea-talk.Rmd
Start of talk for EA society next Thursday

---
title: How it's been doing a statistics PhD plus some research
subtitle: EA Imperial
author: Adam Howes
institute: Imperial College London
date: February 2022
bibliography: citations.bib
output:
  beamer_presentation:
    latex_engine: pdflatex
    highlight: haddock
    fig_width: 7 
    fig_height: 3
    includes:
        in_header: preamble.tex
---

## Academic background

* Maths undergraduate at Durham, no real ideas about career
  * Worked in a pub first year, Cranfield second year, and Basque Center for Applied Mathematics third year
  * Dissertation about Bayesian optimisation 
* Not enough statistics courses at Durham, so did a masters at Warwick
  * Dissertation about Bayesian models for combining sources of evidence
* StatML CDT (Imperial / Oxford) started the year I was applying
  * Somewhat successor to OxWaSP (Oxford / Warwick)

## EA background

* Heard about 80,000 Hours in 3rd year from a friend
* Listened to a few podcasts, found them compelling, first time anything career related really resonated with me
* Read a few books over summer, then started going to the EA society at Warwick
  * Events like retreats, more people in the community, helped to add depth
* First EAG I went to was the year I started at Imperial
  * Volunteering can be a good way to get a place

## Any tips about applying for PhDs

* Find (and call if possible) any previous students of the supervisor
  * What are they doing now?
* Assess the vibe among current graduate students
* Be interested in things, read papers, attend reading groups
  * "List of things you don't need permission to do"
* Doing research as an undergraduate is a good test

## Reflections on PhD experience

* Many (all?) of my friends doing PhDs have had periods where it has negatively effected them
  * "The well-being costs of graduate school versus STIs"
* Little structure, often sporadic feedback, uncertainty about research direction and progress, making a narrow contribution (that perhaps few people care about) to a large field
* Different to undergraduate: not learning material then passing test for another 3/4 years

## What's my PhD about

* HIV/AIDS is an infectious disease which is endemic in sub-Saharan Africa
* Epidemic indicators can be used to guide policy response
* My area of research is Bayesian spatio-temporal statistics, the method we use to perform small-area estimation
  * I'll explain what these terms mean

## Bayesian statistics

## Spatio-temporal statistics

## Career plans going forward