### 2022-02-01 src/docs_03-02-22-ea-talk/03-02-22-ea-talk.Rmd
Finish draft of these slides

---
title: How it's been doing a statistics PhD plus some research
subtitle: EA Imperial
author: Adam Howes
institute: Imperial College London
date: February 2022
bibliography: citations.bib
output:
  beamer_presentation:
    latex_engine: pdflatex
    highlight: haddock
    fig_width: 7 
    fig_height: 3
    includes:
        in_header: preamble.tex
---

## Academic background

* Maths undergraduate at Durham, no real ideas about career
  * Worked in a pub first year, Cranfield second year, and Basque Center for Applied Mathematics third year
  * Dissertation about Bayesian optimisation 
* Not enough statistics courses at Durham, so did a masters at Warwick
  * Dissertation about Bayesian models for combining sources of evidence
* StatML CDT (Imperial / Oxford) started the year I was applying
  * Somewhat successor to OxWaSP (Oxford / Warwick)

## EA background

* Heard about 80,000 Hours in 3rd year from a friend
* Listened to a few podcasts, found them compelling, first time anything career related really resonated with me
* Read a few books over summer, then started going to the EA society at Warwick
  * Events like retreats, more people in the community, helped to add depth
* First EAG I went to was the year I started at Imperial
  * Volunteering can be a good way to get a place

## Is it worth doing a PhD?

* Skills or talent gap for many causes in EA
  * Rather than funding gap
* Independent research skills developed during a PhD important for many causes
  * AI, biorisk, global priorities, global health and development
* Can be useful for signalling expertise
  * Then transfer into e.g. policy
* 3/4+ years, possibly large opportunity cost

## Any tips about applying for PhDs

* Find (and call if possible) any previous students of the supervisor
  * What are they doing now?
* Assess the vibe among current graduate students
* Be interested in things, read papers, attend reading groups
  * "List of things you don't need permission to do"
* Doing research as an undergraduate is a good test

## Some negative reflections on PhD experience

* I, along with many (all?) of my friends doing PhDs, have had periods where it has negatively effected wellbeing
  * "Grad school is worse for public health than STDs" https://www.benkuhn.net/grad/
* Little structure, often sporadic feedback, uncertainty about research direction and progress, narrow contribution (that perhaps few people care about) to a large field, low pay, alignment with supervisor and program, (in some fields) somewhat pyramid scheme-y
* Different to undergraduate: not learning material then passing test for another 3/4 years

## What's my PhD about

* HIV/AIDS is an infectious disease which is endemic in sub-Saharan Africa
* Epidemic indicators can be used to guide policy response
  * Prevalence (how many people are infected)
  * Incidence (how many people newly infected)
  * Coverage of treatments (how many people have (don't have) treatment)
* My area of research is Bayesian spatio-temporal statistics, the method we use to perform small-area estimation
  * (I'll explain what these terms mean)

## Bayesian statistics

* First step in Bayesian statistics is to write a probabilistic model for the data
  * Focus is on understanding and modelling scientific phenomenon, rather than e.g. remembering arcane statistical tests
* Second step is to perform inference (which is usually done with an algorithm)
  * Given we observed certain data, what can we learn about the parameters of the model?
* Things might go wrong with the above two steps with the data, probabilistic model, algorithm
  * Should also perform checks, following a "Bayesian workflow"

## Spatio-temporal statistics

* Say that you have a collection of people who have taken a HIV test, and their results are either positive or negative
* If have the location for these people, this might be important information to include in our probabilistic model
* Similarly, the time that they took the test might also be important
* Spatio-temporal statistics is just when we think these two features of the data are important to take into account
* It allows us to answer spatio-temporal questions
  * What's the situation at time $t$ and location $s$?
  * Often important for policy

## My work

* I'm interested in applied and methodological work
  * Applied: focus is on answering a question about the world
  * Methodological: focus is on developing techniques which can be used to answer questions about the world
* The deliniation between these might not be clear sometimes
  * To answer questions about the world you'll often have to develop new methods
  * To develop new (relevant) methods that you should be thinking about which questions they will allow you to answer to think
* Some examples of my work
  * Applied: estimating numbers of young women at risk
  * Methodological: investigating models for spatial structure

## Estimating numbers of young women at risk

* Women 15-29 are disproportionately at risk of HIV
* To efficiently target programs and resources to the most at risk women, we're working on small-area estimates of the number of women in different risk groups

##

\begin{table}
\centering
\begin{tabularx}{\textwidth}{lXlX}
\toprule
Category & Description & Risk ratio \\ 
 \midrule
None & Not sexually active & 0 \\
Low & One cohabiting sexual partner & 1 \\
High & Non-regular sexual partner(s) & 1.72  \\
Very high & Young women from key populations & 13 \\
\bottomrule
\end{tabularx}
\caption{HIV risk categories and risk ratios.}
\label{tab:risk-categories}
\end{table}

##

```{r echo=FALSE}
knitr::include_graphics("depends/available-surveys.pdf")
```

##

```{r echo=FALSE}
knitr::include_graphics("depends/aids-abstract-A.pdf")
```

##

```{r echo=FALSE}
knitr::include_graphics("depends/aids-abstract-B.pdf")
```

##

```{r echo=FALSE}
knitr::include_graphics("depends/model-direct-benefits-h.pdf")
```

## The Besag model

$$
\phi_i \, | \, \bphi_{-i} \sim \N\left(\frac{\sum_{j: j \sim i} \phi_j}{n_{\delta i}}, \frac{1}{n_{\delta i}\tau_\phi}\right)
$$

Mean of $\phi_i$ average of adjacent units and precision is proportional to $n_{\delta i}$

##

```{r echo=FALSE}
knitr::include_graphics("geometry-graph-zwe.pdf")
```

## Impact of this work

* Could be used to guide HIV interventions for women 15-29
* Interventions suggested include 
  * STI-screening and treatment
  * Economic empowerment,
  * Comprehensive sexuality education
  * PrEP
  * PEP

##

```{r echo=FALSE}
knitr::include_graphics("depends/infections-averted-country.pdf")
```

## Some possible routes for the future

* Small-area estimation of health and development indicators
* Outbreak disease surveillance
* Monitoring of the environment (e.g. sea-ice for climate, air-pollution for global health)
* Statistical ecology (as it relates to animal welfare)
* AI safety (likely the more applied FAT-style work than theoretical)
* Predicting (or helping to win) elections
* Meta EA

## Current plan

* Outbreak disease surveillance for reducing GCBRs
* Longtermist issues likely score better on neglectedness and scope
* Biorisk cause area "competitive" with AI risk
  * Better personal fit to biorisk
* Biosecurity reading group this term, Fridays at St Mary's 12-1!

## Thanks!