### 2022-03-15 src/docs_amazon-poster/amazon-poster.Rmd
Small changes to poster after meeting #77

---
main_topsize: 0.2
main_bottomsize: 0.075
poster_height: "48in"
poster_width: "36in"
font_family: "Open Sans"
#ESSENTIALS
title: Spatio-temporal estimates of risk group proportions for adolescent girls and young women across 13 priority countries in sub-Saharan Africa
author:
  - name: '**Adam Howes**'
    affil: 1
    main: true
    twitter: adamhowes
    email: ath19@ic.ac.uk
  - name: "Kathryn A. Risher"
    affil: 3
  - name: "Van Kinh Nguyen"
    affil: 3
  - name: "Oliver Stevens"
    affil: 3
  - name: "Katherine M. Jia"
    affil: 4
  - name: "Tim M. Wolock"
    affil: 1
  - name: "Lycias Zembe"
    affil: 5
  - name: "Ian Wanyeki"
    affil: 5
  - name: "Mary Mahy"
    affil: 5
  - name: "Clemens Benedikt"
    affil: 5
  - name: "Seth R. Flaxman"
    affil: 6
  - name: "Jeffrey W. Eaton"
    affil: 2
affiliation:
  - num: 1
    address: Department of Mathematics, Imperial College London
  - num: 2
    address: MRC Centre for Global Infectious Disease Analysis, School of Public Health, Imperial College London
  - num: 3
    address: Department of Medical Social Sciences, Feinberg School of Medicine, Northwestern University
  - num: 4
    address: Harvard T. H. Chan School of Public Health, Harvard University
  - num: 5
    address: Joint United Nations Programme on HIV/AIDS, Geneva, Switzerland
  - num: 6
    address: Department of Computer Science, University of Oxford
main_findings:
  - "Successfully **targeting** HIV programming using **risk behaviour** would **greatly increase** new infections reached."
logoleft_name: "IMP_ML_W_N.png"
logoright_name: "UNAIDS-reverse-EN.gif"
# Style & formatting
column_numbers: 3
title_textsize: "60px"
author_textsize: "1.17em"
authorextra_textsize: "35px"
affiliation_textsize: "25px"
affiliation_textcol: '#00000080'
caption_fontsize: "20px"
# Main
main_fontfamily: "Open Sans"
# Poster body options
primary_colour: '#38761C'
secondary_colour: '#3b77D8'
accent_colour: "#5d19c4"
body_bgcol: "#ffffff"
body_textsize: "40px"
body_textcol: "#000000"
reference_textsize: "25px"
# Standard options
bibliography: citations.bib
output: 
  posterdown::posterdown_betterport:
    self_contained: false
    pandoc_args: --mathjax
    highlight: espresso
    number_sections: false
---

# Introduction

* Adolescent girls and young women (AGYW) 15-29 continue to **face higher risk** of HIV, and have been identified as a **priority population** for prevention
* The UNAIDS Global AIDS Strategy **recommends differentiating services** for AGYW geographically based on individual risk behaviour
* We used a spatio-temporal model to estimate the proportion of AGYW in **four behavioural risk categories** $k = 1:4$ in 13 priority countries at a district level in the years 1999-2020

|$k$|Category                 | Risk ratio|
|:--|:------------------------|----------:|
|1  |Not sexually active      |          0|
|2  |One cohabiting partner   |          1|
|3  |Non-regular partner(s)   |       1.72|
|4  |Female sex workers (FSW) |         13|

# Methods

* We used 48 household surveys (AIS, BAIS, DHS, PHIA)
* For the categories $k \in \{1, 2, 3^{+}\}$ we used all the data to fit a survey-weighted multinomial logistic regression via the multinomial-Poisson transformation
\begin{align}
    &y_{itak}^\star \sim \text{xPoisson}(\kappa_{itak}), \\
    &\log(\kappa_{itak}) 
    = \underbrace{\theta_{ita}}_{\text{Observation (IID)}}
    + \underbrace{\beta_k}_{\text{Category (IID)}}
    + \underbrace{\alpha_{ak}}_{\text{Age (IID)}} \\
    &+ \underbrace{\zeta_{ck}}_{\text{Country (IID)}} 
    + \underbrace{\phi_{ik}}_{\text{Spatial (Besag)}} 
    + \underbrace{\gamma_{tk}}_{\text{Temporal (AR1)}}
\end{align}
where $p_{itak} = \text{softmax}(\kappa_{ita})_k$
* To differentiate $3^{+}$ we used the 13 surveys with a paid sex question to fit a survey-weighted logistic regression
\begin{align}
    &y_{ia4}^\star \sim \text{xBinomial} \left( y_{ia3} + y_{ia4}, p_{ia4} / (p_{ia3} + p_{ia4}) \right), \\
    &\text{logit}(p_{ia4} / (p_{ia3} + p_{ia4}))
    = \beta_0
    + \beta_\texttt{cfswever} x_c \\
    &+ \underbrace{\alpha_a}_{\text{Age (IID)}}
    + \underbrace{\zeta_c}_{\text{Country (IID)}}
    + \underbrace{\phi_i}_{\text{Spatial (Besag)}}
\end{align}
where $x_c$ is the proportion of men in each country who are clients of FSW [@hodgins2021cfsw]
* We align our estimates for FSW to the UNAIDS Key Populations Atlas at the national-level [@johnston2022deriving]
* Inference performed with integrated nested Laplace approximation [@rue2009approximate] via `R-INLA`

# Results

```{r echo=FALSE, fig.align='center', out.width='100%', fig.cap='Comparison of targeting strategies'}
knitr::include_graphics("depends/infections-averted.png")
```

# Discussion

* Using a **modular approach** allowed integration of all relevant data
* **Spatio-temporal smoothing** overcame high variance from small sample sizes
* Individual behaviour is a **key determinant** of risk and providing prevention services on the basis of behavior would allow many more new infections to be reached, **especially among FSW**
* Geographic delineation: cohabiting more common in the east and non-regular partner(s) more common in the south
* Large numbers of 15-19 in Mozambique already cohabiting

## Limitations

* No adjustment for **reporting bias**
* **Risk category definitions** could be disputed
* No assessment **practicalities or costs** of strategies

## Future work

* Inclusion of **more surveys** e.g. VACS
* Extension to **general population** of adults 15-49

# References

<div id="refs" class="references"></div>