### 2022-05-06 src/docs_12-05-22-mrc-gida/12-05-22-mrc-gida.Rmd
Presentable version of slides incase laptop dies again. Close #112

---
title: Spatio-temporal estimates of HIV risk group proportions for adolescent girls and young women across 13 priority countries in sub-Saharan Africa
subtitle: MRC Centre for Global Infectious Disease Analysis Seminars
author: Adam Howes
institute: Imperial College London
date: May 2022
bibliography: citations.bib
header-includes:
  - \widowpenalties 1 150
output:
  beamer_presentation:
    latex_engine: pdflatex
    highlight: haddock
    fig_width: 7 
    fig_height: 3
    includes:
        in_header: preamble.tex
---

## Background

* In sub-Saharan Africa, adolescent girls and young women (AGYW) aged 15-29 are disproportionately at risk of HIV infection
* This disparity is because of:
  1. Younger age at first sex
  2. Age patterns of sexual mixing
  3. Structural vulnerabilities and power imbalances
  4. Increased susceptibility to HIV infection

## Stratified prevention

* Prevention packages are made up of a core package, available to everyone, and more intensified interventions
* To have the greatest impact, it's important to priorisite intensified interventions to those at highest risk
* The Global AIDS Strategy 2021-2016 proposed stratifying HIV prevention for AGYW based upon
  1. Population-level HIV incidence in the local population
  2. Individual-level sexual risk behaviour

## Scope for our work

### Goals

1. Enable implementation of prevention strategies stratified by incidence and behaviour
2. Assess the benefits of such approaches

### Approach

1. Estimate the proportion of AGYW in four behavioural risk groups at a district level (in 13 countries identified as priority by The Global Fund)
2. Analyze numbers of new infections reached by stratified prevention strategies

## Data

* We used sexual behaviour data from AIS, BAIS, DHS and PHIA household surveys to place respondents into four risk groups:
  1. $k = 1$: Not sexually active
  2. $k = 2$: One cohabiting sexual partner
  3. $k = 3$: Non-regular sexual partner(s)
  4. $k = 4$: Female sex workers
* District-level HIV incidence, prevalence, population size estimates from the Naomi model [@eaton2021naomi]
* Risk ratios from ALPHA network analysis [@slaymaker2020croi] and UNAIDS analysis led by Keith Sabin

##

```{r echo=FALSE, out.width = "400px"}
knitr::include_graphics("depends/category-flowchart.pdf")
```

##

```{r echo=FALSE, out.width = "400px"}
knitr::include_graphics("depends/available-surveys.png")
```

## Two-stage model for risk group proportions

### Stage 1: $k = 1, 2, 3^{+}$

* Multinomial logistic regression model for the proportion of AGYW in the $k = 1, 2, 3^{+}$ risk groups, using all 47 surveys
* Selected model included:
  * Age country effects (IID)
  * Country effects (IID)
  * Correlated spatial effects (ICAR)
  * Correlated temporal effects (AR1) 
* Multinomial-Poisson transformation allowed us to use `R-INLA` for inference

## Two-stage model for risk group proportions

### Stage 2: $k = 3, 4$

* Logistic regression model for the proportion of those in the $k = 3^{+} = \{3, 4\}$ risk groups who are in the $k = 4$ risk group, using only the 13 surveys with a specific transactional sex question
* Selected model included:
  * Age country effects (IID)
  * Country effects (IID)
  * Correlated spatial effects (ICAR)
  * Clients of FSW covariates [@hodgins2022population]
* Also used `R-INLA` for inference here

## Two-stage model for risk group proportions

### Combination and FSW adjustment

* Take 1000 samples from each model, then multiply suitably to generate estimates for all four risk groups
* We adjusted the $k = 4$ risk group to match national FSW estimates from @johnston2022deriving

\begin{center}
\begin{tcolorbox}[width=0.9\textwidth, colframe={title}, colback={white}, title={}]
$\implies$ Estimates of risk group proportions $\rho_{itak}$ by district, year and age group
\end{tcolorbox}
\end{center}

##

```{r echo=FALSE, out.width="350px", fig.cap = "We found a geographic discontinuity in behaviour between Southern and Eastern Africa."}
knitr::include_graphics("depends/aids-abstract-A.pdf")
```

##

```{r echo=FALSE, out.width="350px", fig.cap = "Here is another view of the discontinuity."}
knitr::include_graphics("depends/aids-abstract-B.pdf")
```

## Benefits of our modelled risk group estimates

* Integration of all relevant surveys
  * Modular approach allowed estimating FSW proportion even for surveys without a specific transactional sex question
* Alleivating small-sample sizes (that caused direct estimates to display large variance)
  * We borrowed information across space, between countries and over surveys so that our estimates more plausibly reflect spatial heterogeneity
* Estimates where there isn't direct data
  * Although some people think of this as "making up data", the data almost never "speak for themselves" (everything is a model)
  * Uncertainty should be higher in regions with infilling: important to transparently communicate this

##

```{r echo=FALSE, out.width = "350px", fig.cap = "Illustration of the problem with direct survey estimates."}
knitr::include_graphics("depends/model-direct-benefits.png")
```

## HIV incidence by risk group

* We used our risk group proportion estimates and relative risk ratio estimates from the literature to disaggregate Naomi general population HIV incidence estimates
\begin{align*}
I_{ia} &= \sum_k \lambda_{iak} N_{iak} \\
&= \sum_k \lambda_{ia2} \text{RR}_k N_{iak}.
\end{align*}

\begin{center}
\begin{tcolorbox}[width=0.9\textwidth, colframe={title}, colback={white}, title={}]
$\implies$ Estimates of HIV incidence $\lambda_{iak}$ and number of new HIV infections $I_{iak}$ by district, age group and risk group
\end{tcolorbox}
\end{center}

## Prioritisation with risk group information

* Suppose we have all of the information (district, age, and risk group)
* Which are the strata with highest incidence?

```{r include=FALSE}
df <- readr::read_csv("depends/incidence-district-sexbehav.csv")

df_area_age_behav <- df %>%
  select(area_id, age_group, population_nosex12m, incidence_nosex12m, starts_with("population_sex"), starts_with("incidence_sex")) %>%
  pivot_longer(
    cols = starts_with("population") | starts_with("incidence"),
    names_to = c("indicator", "category"),
    names_sep = "_"
  ) %>%
  pivot_wider(
    id_cols = c("area_id", "age_group", "indicator", "category"),
    names_from = "indicator",
    values_from = "value"
  )
```

```{r echo=FALSE}
df_area_age_behav %>%
  arrange(desc(incidence)) %>%
  head() %>%
  knitr::kable(digits = 2)
```

## Prioritisation without risk group information

* What about if we lost the risk group information? Now what are the strata with the highest incidence?

```{r include=FALSE}
df_area_age <- df_area_age_behav %>%
  group_by(area_id, age_group) %>%
  summarise(
    incidence = sum(incidence * population, na.rm = TRUE) / sum(population, na.rm = TRUE),
    population = sum(population, na.rm = TRUE)
  ) %>%
  ungroup()
```

```{r echo=FALSE}
df_area_age %>%
  select(area_id, age_group, population, incidence) %>%
  arrange(desc(incidence)) %>%
  head() %>%
  knitr::kable(digits = 2)
```

##

```{r echo=FALSE, out.width = "400px"}
knitr::include_graphics("depends/infections-reached.png")
```

## Limitations

* Simplistic infections reached analysis
  * No accounting for difficulty in reaching each strata
  * Variable intervention effectiveness
  * Change in strata membership
* Under-reporting of high risk sexual behaviours
  * Variation in under-reporting (likely by age, foremost, and location, less so) particularly concerning
* Simple risk groups, not clearly linked to data
  * Didn't consider other important characteristics that may determine risk e.g. condom usage
* Only focused on AGYW 15-29
  * Could be extended to adults of both sexes aged 15-49

## Policy takeaways

* Our risk group estimates can be used to help countries implement the Global AIDS Strategy (user guide currently under development)
* It's important to reach FSW (not a new message!)
* Countries have different epidemic profiles

## Methods takeaways

* `R-INLA` and the multinomial-Poisson trick are a good choice for spatio-temporal multinomial models
  * The `group` and `replicate` features can be used to define quite complex Kronecker product effects
  * However, can run into limitations on model complexity and I might be moving to something more flexible e.g. `tmb`, `aghq` to continue development of this project

## Thanks for listening!

* This project was joint work with members of the HIV inference group, particularly Katie Risher and Jeff Eaton
  * \faHome $\,$ `hiv-inference.org`
* You can find me via
  * \faEnvelope $\,$ `ath19@ic.ac.uk`
  * \faHome $\,$ `athowes.github.io`

## Bibliography {.allowframebreaks}