### 2022-06-17 src/docs_paper/tables-figures.Rmd
Variance proportions, prevalence plots into appendix B. Changes to captions #128

---
title: Supplementary tables and figures for "Spatio-temporal estimates of HIV risk group proportions for adolescent
  girls and young women across 13 priority countries in sub-Saharan Africa"
output:
  pdf_document:
    toc: true
    keep_tex: true
    includes:
      in_header: preamble.tex
bibliography: citations.bib
---

\clearpage

<!-- So that the tables and figures in this appendix are prefixed by "B." -->
\setcounter{table}{0}  
\renewcommand{\thetable}{B.\arabic{table}} 
\setcounter{figure}{0} 
\renewcommand{\thefigure}{B.\arabic{figure}}

# The Global AIDS Strategy

\begin{table}[h]
\centering
\begin{tabularx}{\textwidth}{lX}
\toprule
Prioritisation strata & Criterion \\
\midrule
Low & 0.3-1.0\% incidence and low-risk behaviour, or $<$0.3\% incidence and high-risk behaviour \\
Moderate & 1.0-3.0\% incidence and low-risk behaviour, or 0.3-1.0\% incidence and high-risk behaviour \\
High & 1.0-3.0\% incidence and high-risk behaviour \\
Very high & $>$3.0\% incidence \\
\bottomrule
\end{tabularx}
\caption{Prioritisation strata.}
\label{tab:unaids-strategy-prioritisation}
\end{table}

\begin{table}[h]
\centering
\begin{tabularx}{\textwidth}{lX}
\toprule
Intervention & Details \\
\midrule
STI screening and treatment & (Remain to be filled in: describe targets and commitments in terms of each prioritisation strata) \\
PrEP use & (As above) \\
Access to PEP & (As above) \\
Comprehensive sexuality education & (As above) \\
Economic empowerment & (As above) \\
\bottomrule
\end{tabularx}
\caption{Targets and commitments.}
\label{tab:unaids-strategy-targets}
\end{table}

\clearpage

# Household survey data

## Included surveys

\begin{longtable}{rlrcrrrr}
\toprule
\multicolumn{1}{l}{} &  &  &  & \multicolumn{4}{c}{Sample size} \\ 
\cmidrule(lr){5-8}
\multicolumn{1}{l}{} & Type & Year & Transactional sex question & 15-19 & 20-24 & 25-29 & Total \\ 
\midrule
\multicolumn{1}{l}{Botswana} \\ 
\midrule
 & BAIS & 2013 & \cmark & 557 & 588 & 649 & 1794 \\ 
Total &  &  &  & $557$ & $588$ & $649$ & $1794$ \\ 
\midrule
\multicolumn{1}{l}{Cameroon} \\ 
\midrule
 & DHS & 2004 & \xmark & 2675 & 2207 & 1732 & 6614 \\ 
 & DHS & 2011 & \xmark & 3588 & 3115 & 2655 & 9358 \\ 
 & PHIA & 2017 & \xmark & 2620 & 2339 & 2259 & 7218 \\ 
 & DHS & 2018 & \cmark & 3349 & 2463 & 2345 & 8157 \\ 
Total &  &  &  & $12232$ & $10124$ & $8991$ & $31347$ \\ 
\midrule
\multicolumn{1}{l}{Kenya} \\ 
\midrule
 & DHS & 2003 & \xmark & 1819 & 1709 & 1391 & 4919 \\ 
 & DHS & 2008 & \xmark & 1767 & 1743 & 1419 & 4929 \\ 
 & DHS & 2014 & \xmark & 2861 & 2534 & 2858 & 8253 \\ 
Total &  &  &  & $6447$ & $5986$ & $5668$ & $18101$ \\ 
\midrule
\multicolumn{1}{l}{Lesotho} \\ 
\midrule
 & DHS & 2004 & \xmark & 1761 & 1455 & 1026 & 4242 \\ 
 & DHS & 2009 & \xmark & 1833 & 1543 & 1194 & 4570 \\ 
 & DHS & 2014 & \xmark & 1537 & 1292 & 1067 & 3896 \\ 
 & PHIA & 2017 & \cmark & 1156 & 1202 & 1054 & 3412 \\ 
Total &  &  &  & $6287$ & $5492$ & $4341$ & $16120$ \\ 
\midrule
\multicolumn{1}{l}{Mozambique} \\ 
\midrule
 & AIS & 2009 & \xmark & 1031 & 1106 & 987 & 3124 \\ 
 & DHS & 2011 & \xmark & 2932 & 2299 & 2206 & 7437 \\ 
 & AIS & 2015 & \xmark & 1552 & 1389 & 1080 & 4021 \\ 
Total &  &  &  & $5515$ & $4794$ & $4273$ & $14582$ \\ 
\midrule
\multicolumn{1}{l}{Malawi} \\ 
\midrule
 & DHS & 2000 & \xmark & 2914 & 2998 & 2358 & 8270 \\ 
 & DHS & 2004 & \xmark & 2407 & 2823 & 2135 & 7365 \\ 
 & DHS & 2010 & \xmark & 5031 & 4387 & 4309 & 13727 \\ 
 & DHS & 2015 & \cmark & 5273 & 5094 & 3976 & 14343 \\ 
 & PHIA & 2016 & \cmark & 1646 & 1934 & 1511 & 5091 \\ 
Total &  &  &  & $17271$ & $17236$ & $14289$ & $48796$ \\ 
\midrule
\multicolumn{1}{l}{Namibia} \\ 
\midrule
 & DHS & 2000 & \xmark & 1427 & 1313 & 1098 & 3838 \\ 
 & DHS & 2006 & \xmark & 2203 & 1869 & 1544 & 5616 \\ 
 & DHS & 2013 & \xmark & 1852 & 1709 & 1481 & 5042 \\ 
 & PHIA & 2017 & \cmark & 1491 & 1525 & 1370 & 4386 \\ 
Total &  &  &  & $6973$ & $6416$ & $5493$ & $18882$ \\ 
\midrule
\multicolumn{1}{l}{Eswatini} \\ 
\midrule
 & DHS & 2006 & \xmark & 1265 & 1027 & 731 & 3023 \\ 
 & PHIA & 2017 & \xmark & 1031 & 895 & 811 & 2737 \\ 
Total &  &  &  & $2296$ & $1922$ & $1542$ & $5760$ \\ 
\midrule
\multicolumn{1}{l}{Tanzania} \\ 
\midrule
 & AIS & 2003 & \xmark & 1466 & 1377 & 1270 & 4113 \\ 
 & AIS & 2007 & \xmark & 2137 & 1676 & 1509 & 5322 \\ 
 & DHS & 2010 & \xmark & 2221 & 1860 & 1613 & 5694 \\ 
 & AIS & 2012 & \xmark & 2474 & 1923 & 1815 & 6212 \\ 
 & PHIA & 2016 & \cmark & 2999 & 2845 & 2521 & 8365 \\ 
Total &  &  &  & $11297$ & $9681$ & $8728$ & $29706$ \\ 
\midrule
\multicolumn{1}{l}{Uganda} \\ 
\midrule
 & DHS & 2000 & \xmark & 1687 & 1541 & 1326 & 4554 \\ 
 & DHS & 2006 & \xmark & 1948 & 1660 & 1404 & 5012 \\ 
 & AIS & 2011 & \xmark & 2451 & 2164 & 1921 & 6536 \\ 
 & DHS & 2011 & \xmark & 2025 & 1664 & 1614 & 5303 \\ 
 & DHS & 2016 & \cmark & 4276 & 3782 & 3014 & 11072 \\ 
 & PHIA & 2016 & \xmark & 3289 & 3059 & 2574 & 8922 \\ 
\midrule 
Total &  &  &  & $15676$ & $13870$ & $11853$ & $41399$ \\ 
\midrule
\multicolumn{1}{l}{South Africa} \\ 
\midrule
 & DHS & 2016 & \cmark & 1505 & 1408 & 1397 & 4310 \\ 
Total &  &  &  & $1505$ & $1408$ & $1397$ & $4310$ \\ 
\midrule
\multicolumn{1}{l}{Zambia} \\ 
\midrule
 & DHS & 2007 & \xmark & 1598 & 1405 & 1373 & 4376 \\ 
 & DHS & 2013 & \xmark & 3685 & 3036 & 2789 & 9510 \\ 
 & PHIA & 2016 & \cmark & 2120 & 2045 & 1619 & 5784 \\ 
 & DHS & 2018 & \cmark & 3112 & 2687 & 2166 & 7965 \\ 
Total &  &  &  & $10515$ & $9173$ & $7947$ & $27635$ \\ 
\midrule
\multicolumn{1}{l}{Zimbabwe} \\ 
\midrule
 & DHS & 1999 & \xmark & 1467 & 1230 & 1011 & 3708 \\ 
 & DHS & 2005 & \xmark & 2128 & 1943 & 1438 & 5509 \\ 
 & DHS & 2010 & \xmark & 1963 & 1796 & 1679 & 5438 \\ 
 & DHS & 2015 & \cmark & 2154 & 1777 & 1646 & 5577 \\ 
 & PHIA & 2016 & \cmark & 2114 & 1817 & 1573 & 5504 \\ 
Total &  &  &  & $9826$ & $8563$ & $7347$ & $25736$ \\ 
\midrule 
\midrule 
Total &  &  &  & $106397$ & $95253$ & $82518$ & $284168$ \\ 
\bottomrule
\caption{All of the surveys that we used in our analysis and their sample sizes, disaggregated by respondent age.}
\label{tab:surveys-used}
\end{longtable}

\begin{table}[h]
\centering
\begin{tabularx}{\textwidth}{lX}
\toprule
Survey & Exclusion reason \\ 
 \midrule
MOZ2003DHS & No GPS coordinates available to place survey clusters within districts. \\
TZA2015DHS & Insufficient sexual behaviour questions. \\
UGA2004AIS & Unable to download region boundaries. \\
ZMB2002DHS & No GPS coordinates available to place survey clusters within districts. \\
\bottomrule
\end{tabularx}
\label{tab:surveys-excluded}
\caption{All of that surveys that were excluded from our analysis.}
\end{table}

\clearpage

## Spatial analysis levels

\begin{table}[h]
\centering
\begin{tabularx}{\textwidth}{lXX}
\toprule
Country & Number of areas & Analysis level \\ 
\midrule
Botswana & 27 & 3 \\ 
Cameroon & 58 & 2 \\ 
Kenya & 47 & 2 \\ 
Lesotho & 10 & 1 \\ 
Mozambique & 161 & 3 \\ 
Malawi & 33 & 5 \\ 
Namibia & 38 & 2 \\ 
Eswatini & 4 & 1 \\ 
Tanzania & 195 & 4 \\ 
Uganda & 136 & 3 \\ 
South Africa & 52 & 2 \\ 
Zambia & 116 & 2 \\ 
Zimbabwe & 63 & 2 \\ 
\bottomrule
\end{tabularx}
\label{tab:area-levels}
\caption{The numer of areas and analysis levels for each country that were used in our analysis.}
\end{table}

\clearpage

## Survey questions

\begin{table}[h]
\centering
\begin{tabularx}{\textwidth}{lX}
\toprule
Variable(s) & Description \\
\midrule
\texttt{v501} & Current marital status of the respondent. \\
\texttt{v529} & Computed time since last sexual intercourse. \\
\texttt{v531} & Age at first sexual intercourse--imputed. \\
\texttt{v766b} & Number of sexual partners during the last 12 months (including husband). \\
\texttt{v767[a, b, c]} & Relationship with last three sexual partners. Options are: spouse, boyfriend not living with respondent, other friend, casual acquaintance, relative, commercial sex worker, live-in partner, other. \\
\texttt{v791a} & Had sex in return for gifts, cash or anything else in the past 12 months. Asked only to women 15-24 who are not in a union. \\
\bottomrule
\end{tabularx}
\caption{AIS, BAIS and DHS survey questions.}
\end{table}

\begin{table}[h]
\centering
\begin{tabularx}{\textwidth}{lX}
\toprule
Variable(s) & Description \\
\midrule
\texttt{part12monum} & Number of sexual partners during the last 12 months (including husband). \\
\texttt{part12modkr} & Reason for leaving \texttt{part12monum} blank. \\
\texttt{partlivew[1, 2, 3]} & Does the person you had sex with live in this household? \\
\texttt{partrelation[1, 2, 3]} & Relationship with last three sexual partners. Options are: husband, live-in partner, partner (not living with), ex-spouse/partner, friend/acquaintance, sex worker, sex worker client, stranger, other, don't know, refused. \\
\texttt{sellsx12mo} & Had sex for money and/or gifts in the last 12 months. \\
\texttt{buysx12mo} & Paid money or given gifts for sex in the last 12 months. \\
\bottomrule
\end{tabularx}
\caption{PHIA survey questions.}
\end{table}

## Risk group allocation

\newpage

\begin{figure}[h!]
    \centering
    \includegraphics{depends/category-flowchart.pdf}
    \caption{Flowchart describing allocation of respondents to risk groups.}
    \label{fig:category-flowchart}
\end{figure}

\clearpage

# Miscellaneous figures

<!-- Might find a home for these. -->

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics{depends/model-direct-benefits.pdf}}
\caption{Illustration our model results for AGYW 20-24 in Tanzania in 2010 in the cohabiting risk group. Compared to the direct survey results, our spatio-temporally smoothed estimates more plausibly represent district-level heterogeneity, as well as imputing any districts with missing data.}
\label{fig:model-direct-benefits}
\end{figure}

\begin{figure}[h]
    \centering
    \makebox[\textwidth][c]{\includegraphics{depends/aaa-variance-proportions.pdf}}
    \caption{Proportion of variance explained by each random effect (Sobol' indices) when the multinomial regression model is fit to each country individually. In this setting, country-category random effects are not included in the model and year-category random effects are replaced by survey-category random effects (for countries with surveys in multiple years). Countries are ordered by the proportion of their variance which is explained by the area-category random effects.}
\label{fig:aaa-variance-proportions}
\end{figure}

\clearpage

# Country-specific figures

## Comparison of direct and modelled risk group estimates

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=1]{depends/temporal-interpolation-ribbon.pdf}}
\caption{Comparison of modelled and direct national-level estimates in 1999-2018 in Botswana. Estimates are described as "partially direct" when there are no surveys containing a transactional sex question in a country-age-group and we instead used modelled logistic regression estimates to differentiate the direct estimates.}
\label{fig:temporal-interpolation-ribbon-1}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=2]{depends/temporal-interpolation-ribbon.pdf}}
\caption{Comparison of modelled and direct national-level estimates in 1999-2018 in Cameroon. Estimates are described as "partially direct" when there are no surveys containing a transactional sex question in a country-age-group and we instead used modelled logistic regression estimates to differentiate the direct estimates.}
\label{fig:temporal-interpolation-ribbon-2}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=3]{depends/temporal-interpolation-ribbon.pdf}}
\caption{Comparison of modelled and direct national-level estimates in 1999-2018 in Kenya. Estimates are described as "partially direct" when there are no surveys containing a transactional sex question in a country-age-group and we instead used modelled logistic regression estimates to differentiate the direct estimates.}
\label{fig:temporal-interpolation-ribbon-3}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=4]{depends/temporal-interpolation-ribbon.pdf}}
\caption{Comparison of modelled and direct national-level estimates in 1999-2018 in Lesotho. Estimates are described as "partially direct" when there are no surveys containing a transactional sex question in a country-age-group and we instead used modelled logistic regression estimates to differentiate the direct estimates.}
\label{fig:temporal-interpolation-ribbon-4}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=5]{depends/temporal-interpolation-ribbon.pdf}}
\caption{Comparison of modelled and direct national-level estimates in 1999-2018 in Mozambique. Estimates are described as "partially direct" when there are no surveys containing a transactional sex question in a country-age-group and we instead used modelled logistic regression estimates to differentiate the direct estimates.}
\label{fig:temporal-interpolation-ribbon-5}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=6]{depends/temporal-interpolation-ribbon.pdf}}
\caption{Comparison of modelled and direct national-level estimates in 1999-2018 in Malawi. Estimates are described as "partially direct" when there are no surveys containing a transactional sex question in a country-age-group and we instead used modelled logistic regression estimates to differentiate the direct estimates.}
\label{fig:temporal-interpolation-ribbon-6}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=7]{depends/temporal-interpolation-ribbon.pdf}}
\caption{Comparison of modelled and direct national-level estimates in 1999-2018 in Namibia. Estimates are described as "partially direct" when there are no surveys containing a transactional sex question in a country-age-group and we instead used modelled logistic regression estimates to differentiate the direct estimates.}
\label{fig:temporal-interpolation-ribbon-7}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=8]{depends/temporal-interpolation-ribbon.pdf}}
\caption{Comparison of modelled and direct national-level estimates in 1999-2018 in Eswatini. Estimates are described as "partially direct" when there are no surveys containing a transactional sex question in a country-age-group and we instead used modelled logistic regression estimates to differentiate the direct estimates.}
\label{fig:temporal-interpolation-ribbon-8}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=9]{depends/temporal-interpolation-ribbon.pdf}}
\caption{Comparison of modelled and direct national-level estimates in 1999-2018 in Tanzania. Estimates are described as "partially direct" when there are no surveys containing a transactional sex question in a country-age-group and we instead used modelled logistic regression estimates to differentiate the direct estimates.}
\label{fig:temporal-interpolation-ribbon-9}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=10]{depends/temporal-interpolation-ribbon.pdf}}
\caption{Comparison of modelled and direct national-level estimates in 1999-2018 in Uganda. Estimates are described as "partially direct" when there are no surveys containing a transactional sex question in a country-age-group and we instead used modelled logistic regression estimates to differentiate the direct estimates.}
\label{fig:temporal-interpolation-ribbon-10}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=11]{depends/temporal-interpolation-ribbon.pdf}}
\caption{Comparison of modelled and direct national-level estimates in 1999-2018 in South Africa. Estimates are described as "partially direct" when there are no surveys containing a transactional sex question in a country-age-group and we instead used modelled logistic regression estimates to differentiate the direct estimates.}
\label{fig:temporal-interpolation-ribbon-11}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=12]{depends/temporal-interpolation-ribbon.pdf}}
\caption{Comparison of modelled and direct national-level estimates in 1999-2018 in Zambia. Estimates are described as "partially direct" when there are no surveys containing a transactional sex question in a country-age-group and we instead used modelled logistic regression estimates to differentiate the direct estimates.}
\label{fig:temporal-interpolation-ribbon-12}
\end{figure}

<!-- \begin{figure}[h] -->
<!-- \centering -->
<!-- \makebox[\textwidth][c]{\includegraphics[page=13]{depends/temporal-interpolation-ribbon.pdf}} -->
<!-- \caption{Comparison of modelled and direct national-level estimates in 1999-2018 in Zimbabwe. Estimates are described as "partially direct" when there are no surveys containing a transactional sex question in a country-age-group and we instead used modelled logistic regression estimates to differentiate the direct estimates.} -->
<!-- \label{fig:temporal-interpolation-ribbon-13} -->
<!-- \end{figure} -->

\clearpage

## HIV prevalence

\begin{figure}[h]
    \centering
    \makebox[\textwidth][c]{\includegraphics[page=1]{depends/prev-district-sexbehav-logit.pdf}}
    \caption{District-level HIV prevalence for each of the risk groups in 2018 in Botswana.}
    \label{fig:prevalence-district-sexbehav-1}
\end{figure}

\begin{figure}[h]
    \centering
    \makebox[\textwidth][c]{\includegraphics[page=2]{depends/prev-district-sexbehav-logit.pdf}}
    \caption{District-level HIV prevalence for each of the risk groups in 2018 in Cameroon.}
    \label{fig:prevalence-district-sexbehav-2}
\end{figure}

\begin{figure}[h]
    \centering
    \makebox[\textwidth][c]{\includegraphics[page=3]{depends/prev-district-sexbehav-logit.pdf}}
    \caption{District-level HIV prevalence for each of the risk groups in 2018 in Kenya.}
    \label{fig:prevalence-district-sexbehav-3}
\end{figure}

\begin{figure}[h]
    \centering
    \makebox[\textwidth][c]{\includegraphics[page=4]{depends/prev-district-sexbehav-logit.pdf}}
    \caption{District-level HIV prevalence for each of the risk groups in 2018 in Lesotho.}
    \label{fig:prevalence-district-sexbehav-4}
\end{figure}

\begin{figure}[h]
    \centering
    \makebox[\textwidth][c]{\includegraphics[page=5]{depends/prev-district-sexbehav-logit.pdf}}
    \caption{District-level HIV prevalence for each of the risk groups in 2018 in Mozambique.}
    \label{fig:prevalence-district-sexbehav-5}
\end{figure}

\begin{figure}[h]
    \centering
    \makebox[\textwidth][c]{\includegraphics[page=6]{depends/prev-district-sexbehav-logit.pdf}}
    \caption{District-level HIV prevalence for each of the risk groups in 2018 in Malawi.}
    \label{fig:prevalence-district-sexbehav-6}
\end{figure}

\begin{figure}[h]
    \centering
    \makebox[\textwidth][c]{\includegraphics[page=7]{depends/prev-district-sexbehav-logit.pdf}}
    \caption{District-level HIV prevalence for each of the risk groups in 2018 in Namibia.}
    \label{fig:prevalence-district-sexbehav-7}
\end{figure}

\begin{figure}[h]
    \centering
    \makebox[\textwidth][c]{\includegraphics[page=8]{depends/prev-district-sexbehav-logit.pdf}}
    \caption{District-level HIV prevalence for each of the risk groups in 2018 in Eswatini.}
    \label{fig:prevalence-district-sexbehav-8}
\end{figure}

\begin{figure}[h]
    \centering
    \makebox[\textwidth][c]{\includegraphics[page=9]{depends/prev-district-sexbehav-logit.pdf}}
    \caption{District-level HIV prevalence for each of the risk groups in 2018 in Tanzania.}
    \label{fig:prevalence-district-sexbehav-9}
\end{figure}

\begin{figure}[h]
    \centering
    \makebox[\textwidth][c]{\includegraphics[page=10]{depends/prev-district-sexbehav-logit.pdf}}
    \caption{District-level HIV prevalence for each of the risk groups in 2018 in Uganda.}
    \label{fig:prevalence-district-sexbehav-10}
\end{figure}

\begin{figure}[h]
    \centering
    \makebox[\textwidth][c]{\includegraphics[page=11]{depends/prev-district-sexbehav-logit.pdf}}
    \caption{District-level HIV prevalence for each of the risk groups in 2018 in South Africa.}
    \label{fig:prevalence-district-sexbehav-11}
\end{figure}

\begin{figure}[h]
    \centering
    \makebox[\textwidth][c]{\includegraphics[page=12]{depends/prev-district-sexbehav-logit.pdf}}
    \caption{District-level HIV prevalence for each of the risk groups in 2018 in Zambia.}
    \label{fig:prevalence-district-sexbehav-12}
\end{figure}

\begin{figure}[h]
    \centering
    \makebox[\textwidth][c]{\includegraphics[page=13]{depends/prev-district-sexbehav-logit.pdf}}
    \caption{District-level HIV prevalence for each of the risk groups in 2018 in Zimbabwe.}
    \label{fig:prevalence-district-sexbehav-13}
\end{figure}

\clearpage

## HIV incidence

\begin{figure}[h]
    \centering
    \makebox[\textwidth][c]{\includegraphics[page=1]{depends/incidence-district-sexbehav.pdf}}
    \caption{District-level HIV incidence for each of the risk groups in 2018 in Botswana.}
    \label{fig:incidence-district-sexbehav-1}
\end{figure}

\begin{figure}[h]
    \centering
    \makebox[\textwidth][c]{\includegraphics[page=2]{depends/incidence-district-sexbehav.pdf}}
    \caption{District-level HIV incidence for each of the risk groups in 2018 in Cameroon.}
    \label{fig:incidence-district-sexbehav-2}
\end{figure}

\begin{figure}[h]
    \centering
    \makebox[\textwidth][c]{\includegraphics[page=3]{depends/incidence-district-sexbehav.pdf}}
    \caption{District-level HIV incidence for each of the risk groups in 2018 in Kenya.}
    \label{fig:incidence-district-sexbehav-3}
\end{figure}

\begin{figure}[h]
    \centering
    \makebox[\textwidth][c]{\includegraphics[page=4]{depends/incidence-district-sexbehav.pdf}}
    \caption{District-level HIV incidence for each of the risk groups in 2018 in Lesotho.}
    \label{fig:incidence-district-sexbehav-4}
\end{figure}

\begin{figure}[h]
    \centering
    \makebox[\textwidth][c]{\includegraphics[page=5]{depends/incidence-district-sexbehav.pdf}}
    \caption{District-level HIV incidence for each of the risk groups in 2018 in Mozambique.}
    \label{fig:incidence-district-sexbehav-5}
\end{figure}

\begin{figure}[h]
    \centering
    \makebox[\textwidth][c]{\includegraphics[page=6]{depends/incidence-district-sexbehav.pdf}}
    \caption{District-level HIV incidence for each of the risk groups in 2018 in Malawi.}
    \label{fig:incidence-district-sexbehav-6}
\end{figure}

\begin{figure}[h]
    \centering
    \makebox[\textwidth][c]{\includegraphics[page=7]{depends/incidence-district-sexbehav.pdf}}
    \caption{District-level HIV incidence for each of the risk groups in 2018 in Namibia.}
    \label{fig:incidence-district-sexbehav-7}
\end{figure}

\begin{figure}[h]
    \centering
    \makebox[\textwidth][c]{\includegraphics[page=8]{depends/incidence-district-sexbehav.pdf}}
    \caption{District-level HIV incidence for each of the risk groups in 2018 in Eswatini.}
    \label{fig:incidence-district-sexbehav-8}
\end{figure}

\begin{figure}[h]
    \centering
    \makebox[\textwidth][c]{\includegraphics[page=9]{depends/incidence-district-sexbehav.pdf}}
    \caption{District-level HIV incidence for each of the risk groups in 2018 in Tanzania.}
    \label{fig:incidence-district-sexbehav-9}
\end{figure}

\begin{figure}[h]
    \centering
    \makebox[\textwidth][c]{\includegraphics[page=10]{depends/incidence-district-sexbehav.pdf}}
    \caption{District-level HIV incidence for each of the risk groups in 2018 in Uganda.}
    \label{fig:incidence-district-sexbehav-10}
\end{figure}

\begin{figure}[h]
    \centering
    \makebox[\textwidth][c]{\includegraphics[page=11]{depends/incidence-district-sexbehav.pdf}}
    \caption{District-level HIV incidence for each of the risk groups in 2018 in South Africa.}
    \label{fig:incidence-district-sexbehav-11}
\end{figure}

\begin{figure}[h]
    \centering
    \makebox[\textwidth][c]{\includegraphics[page=12]{depends/incidence-district-sexbehav.pdf}}
    \caption{District-level HIV incidence for each of the risk groups in 2018 in Zambia.}
    \label{fig:incidence-district-sexbehav-12}
\end{figure}

\begin{figure}[h]
    \centering
    \makebox[\textwidth][c]{\includegraphics[page=13]{depends/incidence-district-sexbehav.pdf}}
    \caption{District-level HIV incidence for each of the risk groups in 2018 in Zimbabwe.}
    \label{fig:incidence-district-sexbehav-13}
\end{figure}

\clearpage

## Expected new infections reached

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=1]{depends/infections-reached-country.pdf}}
\caption{Percentage of expected new infections reached taking a variety of risk stratification approaches against the percentage of at risk population reached in Botswana.}
\label{fig:infections-reached-country-1}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=2]{depends/infections-reached-country.pdf}}
\caption{Percentage of expected new infections reached taking a variety of risk stratification approaches against the percentage of at risk population reached in Cameroon.}
\label{fig:infections-reached-country-2}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=3]{depends/infections-reached-country.pdf}}
\caption{Percentage of expected new infections reached taking a variety of risk stratification approaches against the percentage of at risk population reached in Kenya.}
\label{fig:infections-reached-country-3}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=4]{depends/infections-reached-country.pdf}}
\caption{Percentage of expected new infections reached taking a variety of risk stratification approaches against the percentage of at risk population reached in Lesotho.}
\label{fig:infections-reached-country-4}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=5]{depends/infections-reached-country.pdf}}
\caption{Percentage of expected new infections reached taking a variety of risk stratification approaches against the percentage of at risk population reached in Mozambique.}
\label{fig:infections-reached-country-5}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=6]{depends/infections-reached-country.pdf}}
\caption{Percentage of expected new infections reached taking a variety of risk stratification approaches against the percentage of at risk population reached in Malawi.}
\label{fig:infections-reached-country-6}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=7]{depends/infections-reached-country.pdf}}
\caption{Percentage of expected new infections reached taking a variety of risk stratification approaches against the percentage of at risk population reached in Namibia.}
\label{fig:infections-reached-country-7}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=8]{depends/infections-reached-country.pdf}}
\caption{Percentage of expected new infections reached taking a variety of risk stratification approaches against the percentage of at risk population reached in Eswatini.}
\label{fig:infections-reached-country-8}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=9]{depends/infections-reached-country.pdf}}
\caption{Percentage of expected new infections reached taking a variety of risk stratification approaches against the percentage of at risk population reached in Tanzania.}
\label{fig:infections-reached-country-9}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=10]{depends/infections-reached-country.pdf}}
\caption{Percentage of expected new infections reached taking a variety of risk stratification approaches against the percentage of at risk population reached in Uganda.}
\label{fig:infections-reached-country-10}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=11]{depends/infections-reached-country.pdf}}
\caption{Percentage of expected new infections reached taking a variety of risk stratification approaches against the percentage of at risk population reached in South Africa.}
\label{fig:infections-reached-country-11}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=12]{depends/infections-reached-country.pdf}}
\caption{Percentage of expected new infections reached taking a variety of risk stratification approaches against the percentage of at risk population reached in Zambia.}
\label{fig:infections-reached-country-12}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=13]{depends/infections-reached-country.pdf}}
\caption{Percentage of expected new infections reached taking a variety of risk stratification approaches against the percentage of at risk population reached in Zimbabwe.}
\label{fig:infections-reached-country-13}
\end{figure}