### 2022-07-09 src/docs_paper/paper.Rmd
Draft of fixing comments from MM #128

---
title: Spatio-temporal estimates of HIV risk group proportions for adolescent girls and young women across 13 priority countries in sub-Saharan Africa
author:
  - name: Adam Howes
    email: ath19@ic.ac.uk
    affiliation: [A, B]
    corresponding: ath19@ic.ac.uk
  - name: Kathryn A. Risher 
    affiliation: [C, B]
  - name: Van Kính Nguyen
    affiliation: B
  - name: Oliver Stevens
    affiliation: B
  - name: Katherine M. Jia
    affiliation: D
  - name: Timothy M. Wolock
    affiliation: [A, B]
  - name: Rachel Esra
    affiliation: B    
  - name: Lycias Zembe
    affiliation: E
  - name: Ian Wanyeki
    affiliation: E
  - name: Mary Mahy
    affiliation: E
  - name: Clemens Benedikt
    affiliation: E
  - name: Seth R. Flaxman
    affiliation: F
  - name: Jeffrey W. Eaton
    affiliation: B
address:
  - code: A
    address: Department of Mathematics, Imperial College London
  - code: B
    address: MRC Centre for Global Infectious Disease Analysis, School of Public Health, Imperial College London
  - code: C
    address: Department of Medical Social Sciences, Feinberg School of Medicine, Northwestern University
  - code: D
    address: Harvard T. H. Chan School of Public Health, Harvard University
  - code: E
    address: Joint United Nations Programme on HIV/AIDS, Geneva, Switzerland
  - code: F
    address: Department of Computer Science, University of Oxford
bibliography: citations.bib
output:
  rticles::plos_article:
    includes:
      in_header: preamble.tex
# plos.csl can be download and used locally
csl: http://www.zotero.org/styles/plos
---

```{r echo = FALSE}
options(scipen = 100)

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  dpi = 320,
  cache = TRUE,
  out.width = "95%",
  fig.align = 'center'
)

comma <- function(x) formatC(x, format = "d", big.mark = ",")
```

```{r}
library(tidyverse)

priority_iso3 <- multi.utils::priority_iso3()
analysis_level <- multi.utils::analysis_level()
admin1_level <- multi.utils::admin1_level()

hiv_estimates <- readxl::read_excel(
  "HIV2021Estimates_Global_PLWHV-NI-Pop_6June2022.xlsx", 
  sheet = "Table"
) %>%
  filter(ISO3 %in% priority_iso3) %>%
  select(-c(`...1`, "UN", "ISO2", `...5`, "Region"))

population_comparison <- hiv_estimates %>%
  select(
    "ISO3",
    "L- Population by age and sex 15-19 ; Female", 
    "L- Population by age and sex 20-24 ; Female", 
    "L- Population by age and sex 25-29 ; Female",
    "L- Population aged 15-49 Male+Female"
  ) %>%
  mutate(pop_female_Y015_029 = `L- Population by age and sex 15-19 ; Female` +
    `L- Population by age and sex 20-24 ; Female` + 
    `L- Population by age and sex 25-29 ; Female`
  ) %>%
  summarise(
    total = sum(`L- Population aged 15-49 Male+Female`),
    subset = sum(pop_female_Y015_029)
  )

population_percent <- 100 * round(population_comparison$subset / population_comparison$total, digits = 2)

infections_comparison <- hiv_estimates %>%
  select(
    "ISO3",
    "L- New infections by age 15-19 ; Female", 
    "L- New infections by age 20-24 ; Female", 
    "L- New infections by age 25-29 ; Female",
    "L- New HIV infections (15-49) Male+Female"
  ) %>%
  mutate(inf_female_Y015_029 = `L- New infections by age 15-19 ; Female` +
    `L- New infections by age 20-24 ; Female` + 
    `L- New infections by age 25-29 ; Female`
  ) %>%
  summarise(
    total = sum(`L- New HIV infections (15-49) Male+Female`),
    subset = sum(inf_female_Y015_029)
  )

infections_percent <- 100 * round(infections_comparison$subset / infections_comparison$total, digits = 2)

inc_comparison <- hiv_estimates %>%
  select(
    "ISO3",
    "L- New infections by age 15-19 ; Female", 
    "L- New infections by age 20-24 ; Female", 
    "L- New infections by age 25-29 ; Female",
    "L- New infections by age 15-19 ; Male", 
    "L- New infections by age 20-24 ; Male", 
    "L- New infections by age 25-29 ; Male",
    "L- Population by age and sex 15-19 ; Female", 
    "L- Population by age and sex 20-24 ; Female", 
    "L- Population by age and sex 25-29 ; Female",
    "L- Population by age and sex 15-19 ; Male", 
    "L- Population by age and sex 20-24 ; Male", 
    "L- Population by age and sex 25-29 ; Male",
  ) %>%
  mutate(
    inf_female_Y015_029 = `L- New infections by age 15-19 ; Female` +
    `L- New infections by age 20-24 ; Female` + 
    `L- New infections by age 25-29 ; Female`,
    inf_male_Y015_029 = `L- New infections by age 15-19 ; Male` +
    `L- New infections by age 20-24 ; Male` + 
    `L- New infections by age 25-29 ; Male`,
    pop_female_Y015_029 = `L- Population by age and sex 15-19 ; Female` +
    `L- Population by age and sex 20-24 ; Female` + 
    `L- Population by age and sex 25-29 ; Female`,
    pop_male_Y015_029 = `L- Population by age and sex 15-19 ; Male` +
    `L- Population by age and sex 20-24 ; Male` + 
    `L- Population by age and sex 25-29 ; Male`
  ) %>%
  summarise(
    inc_female_Y015_029 = sum(inf_female_Y015_029) / sum(pop_female_Y015_029),
    inc_male_Y015_029 = sum(inf_male_Y015_029) / sum(pop_male_Y015_029)
  )

inc_ratio <- round(inc_comparison$inc_female_Y015_029 / inc_comparison$inc_male_Y015_029, digits = 1)
```

\clearpage

# Introduction

<!-- TODO: Follow up with Clemens r.e. FSW group name: -->
<!-- There are a couple of options: female sex workers, young women from key populations (and clarify in the text that you mean sex work), AGYW who sell sex. Many people who do this work all the time use "FSW" so likely fine, but there's also a move away from defining someone by their job making "who sell sex" possibly the preferred option. -->

<!-- IW notes that AGYW typically refers to 15-24 -->
<!-- Feels clunky to me to say AGYW and women 25-29 -->
In sub-Saharan Africa, adolescent girls and young women (AGYW) aged 15-29 are `r population_percent`\% of the population but `r infections_percent`\% of new HIV infections [@unaids2021update].
HIV incidence amongst AGYW is `r inc_ratio` times higher than among similarly aged males, due to structural vulnerability and power imbalances, age patterns of sexual mixing, younger age at first sex, and increased susceptibility to HIV infection [@joint2016hiv].
AGYW have therefore been identified as a priority population for HIV primary prevention, with significant investments being made in prevention programming [@saul2018dreams; @global2018measurement].

The Global AIDS Strategy 2021-2026 [@unaids2021global], adopted by the United Nations (UN) General Assembly in June 2021, proposed stratifying packages of HIV prevention provided to AGYW based on both local population-level HIV incidence and individual-level sexual risk behaviour, to promote more efficient priorisation of prevention services [@anderson2014maximising; @mcgillen2016optimum; @kelly2018global; @jia2022risk].
Four prioritisation strata (Table B.1) were defined based on: (1) subnational annual incidence (<0.3\%, 0.3-1.0\%, 1.0-3.0\% and >3.0\%), and (2) self-reported high-risk behaviour or recent STI infection.
<!-- NB: Check again about the recent STI infection aspect. -->
The strategy encourages programmes to  define targets for the proportion of AGYW to be reached with a range of interventions (Table B.2) based upon this prioritisation strata [@unaids2021global].
All AGYW are recommended to have access to a basic package of HIV prevention, while those with with high risk behaviours in moderate incidence settings and all AGYW in very high incidence settings are recommended to have access to enhanced intervention packages [@joint2016hiv; @birdthistle2018evaluating].
<!-- TODO: Add "basic package of HIV prevention consisting of... -->
These interventions may include STI screening and treatment, access to pre-exposure prophylaxis (PrEP), access to post-exposure prophylaxis (PEP), comprehensive sexuality education, and economic empowerment.

Implementation of a stratified HIV prevention strategy by national HIV programmes and stakeholders requires data on the population size and HIV incidence in each risk group by location.
We developed a multinomial Bayesian spatio-temporal model to estimate the proportion of AGYW aged 15-29 years in four behavioural risk groups, stratified by district, year, and five-year age group.
We focused on 13 countries in sub-Saharan Africa which have been identified by the Global Fund to Fight AIDS, TB, and Malaria [@global2018measurement] as priority countries for implementation of AGYW HIV prevention.
Our methodology is standardised across countries, allowing prioritisation both within and between countries.
<!-- Note: Above sentence borrowed from TOR document. To back it up, needs more between country analysis in results section. -->
We analysed the extent to which the risk group proportions varied across districts, age groups, between countries, and over time.
Using our estimates, we calculated the HIV prevalence, people living with HIV (PLHIV), HIV incidence, and expected number of new HIV infections in each risk group by disaggregating district-level estimates from the Naomi model [@eaton2021naomi].
Finally, we quantified the increased efficiency of HIV prevention, in terms of the expected number of new infections that could be preemptively reached, by stratified prioritisation of interventions by risk group, location and age group.

# Methodology

\begin{table}[htb]
\small
\centering
\begin{tabularx}{\textwidth}{lp{5cm}ll}
\toprule
Risk group & Description & Local HIV incidence & Incidence ratio \\
\midrule
None & Not sexually active & -- & 0.0 \\
Low & One cohabiting partner & -- & 1.0 (Baseline) \\
High & Non-regular or multiple partner(s) & -- & 1.72 \\
Very high & Transactional sex (adjusted to correspond to female sex workers) & $<$0.1\% & 25.0 \\
 & & 0.1-0.3\% & 13.0 \\
 & & 0.3-1.0\% & 9.0 \\
 & & 1.0-3.0\% & 6.0 \\
 & & $>$3.0\% & 3.0 \\
\bottomrule
\end{tabularx}
\caption{HIV risk groups and assumed HIV incidence rate ratio for each risk group relative to AGYW with one cohabiting sexual partner. Among FSW, the incidence rate ratio depended on the level of HIV incidence among the general population. The incidence rate ratio for women with non-regular or multiple sexual partner(s) was derived from analysis of ALPHA network data. Non-regular partners are defined to be non cohabiting. The transactional sex risk group is adjusted during analysis to correspond to female sex worker, and incidence rate ratios among FSW were dervied based on patterns of relative HIV prevalence among FSW compared to general population prevalence. When the local HIV incidence in the general population is higher, the incidence rate ratio for FSW is lower.}
\label{tab:risk-groups}
\small
\end{table}

## Data

We analysed nationally-representative household survey data from 13 countries: Botswana, Cameroon, Kenya, Lesotho, Malawi, Mozambique, Namibia, South Africa, Eswatini, Tanzania, Uganda, Zambia and Zimbabwe.
We included surveys conducted in these countries in the years 1999-2018 in which women were interviewed about their sexual behaviour and sufficient geographic information was available to locate survey clusters to districts.
Demographic and Health Surveys (DHS) [@measure2012sampling], AIDS Indicator Surveys (AIS) [@ais], Population-based HIV Impact Assessment (PHIA) [@phia] surveys, and the Botswana AIDS Impact Survey 2013 (BAIS) [@botswana2013botswana] were included.

For each survey, we classified female respondents aged 15-29 years into one of four behavioural risk groups according to reported sexual risk behaviour in the past 12 months.
These risk groups were: not sexually active, one cohabiting sexual partner, non-regular or multiple sexual partner(s), and AGYW who report transactional sex (Table \ref{tab:risk-groups}).
In the case of inconsistent responses, women were categorised according to the highest risk group they fell into, ensuring that the categories were mutually exclusive.
<!-- TODO: Include example of inconsistent result. -->
Exact survey questions varied slightly across survey types and between survey phases (Appendix B).
Questions captured information about whether the respondent had been sexually active in the past twelve months, and if so how with many partners.
For their three most recent partners, respondents were also asked about the type of partnership (spouse, cohabiting partner, partner not cohabiting with respondent, friend, sex worker, sex work client, and other).

Some surveys included a specific question asking if the respondent had received or given money or gifts for sex in the past twelve months.
In these surveys, 2.64\% of women reported transactional sex.
In surveys without such a question, women almost never (0.01\%) answered that one of their three most recent partners was a sex work client.
Due to this incomparably across surveys, we did not include surveys without a specific transactional sex question when estimating the proportion of the population who engaged in transactional sex.
We focused on estimating the proportion of women who reported transactional sex at a district level, and subsequently adjusted these proportions to align to national estimates for the number of female sex workers.

We used estimates of population, people living with HIV (PLHIV) and new HIV infections stratified by district and age group from HIV estimates published by UNAIDS that were developed using the Naomi model [@eaton2021naomi].
The model synthesises data from multiple sources to produce subnational estimates of indicators of interest, and has been used by countries as a part of the HIV estimates process supported by UNAIDS.
<!-- and so far has been used by X countries could be better (adds some information that it's widely used) -->
The administrative area hierarchy and geographic boundaries we used correspond to those used for health service planning by countries, exceptions being Cameroon and Kenya where we conducted analysis one level higher at the department and county levels, respectively (Table B.5).
We used the most recent 2022 estimates for all countries, apart from Mozambique where, due to data accuracy concerns, we used the 2021 estimates (in which the Cabo Delgado province is excluded due to disruption by conflict).

## Two-stage model for sexual risk group proportions

To estimate the proportion of AGYW in each risk group we took a two-stage modelling approach.
First, we fit a spatio-temporal multinomial model stratified by district, year (1999-2018) and five-year age group (15-19, 20-24 and 25-29) for the proportion of AGYW in three categories: (1) not sexually active, (2) one cohabiting partner, and (3) either non-regular or multiple partner(s), or transactional sex.
Combining the two highest risk groups (high and very high) in this way allowed data from all surveys to be included in this first stage model.
Second, we fit a spatial logistic regression model separating those who have non-regular or multiple partner(s) from those who reported transactional sex, stratified by district, and five-year age group, and using only data from the surveys with a specific transactional sex question.
As surveys were only available in the years 2013-2018, we assumed the proportion in the very high risk group among those in the two highest risk groups was constant over time.
We combined the two models using 1000 samples from each posterior distribution to produce samples for all four risk groups.
Finally, we adjusted the samples from the transactional sex category to match age- and country-specific FSW population size estimates.
We modified the samples from the non-regular or multiple partner(s) risk group to ensure that after adjustment the risk group proportions still summed to one.
FSW population size estimates by age were obtained by disaggregating national 15-49 FSW population size estimates [@stevens2022estimating] using the FSW age distribution in South Africa from the Thembisa model [@johnson2020thembisa] in combination with country-specific age at first sex distributions [@nguyen2022trends].
Further technical details are in Appendix A.1.

We considered eight model specifications for the space-age-time multinomial model for the three risk groups (Appendix A.1.2).
All models included intercepts for each risk group, age, country, and age-country effects, a spatial effect allowing variation by district, and a temporal effect allowing variation over time.
Alternative models varied the specification for the spatial effect between independent spatial effects versus spatially correlated effects and independent temporal effects versus first-order auto-regressive temporal effects.
Each alternative was considered with an additional spatio-temporal interaction term allowing the temporal trend to vary across districts.
To understand the importance of each part of the model we analysed the relative sizes of the variance parameters for each effect.

For the logistic regression model of the proportion engaging in transactional sex among those with non-regular or multiple partner(s), we considered six specifications (Appendix A.1.3).
Each included an intercept, age and country effects, and a spatial effect allowing district-level variation.
Both independent and spatially correlated spatial effects were considered.
To improve estimation with sparse data, during model selection we considered alternatives with national-level covariates for either the proportion of men who reported ever having paid for sex or having paid for sex in the last twelve months [@hodgins2022population].

We performed inference using integrated nested Laplace approximation [@rue2009approximate] via the `R-INLA` package [@martins2013bayesian].
It is not possible to directly fit multinomial logistic regression models in `R-INLA`, so we used the multinomial-Poisson transformation [@baker1994multinomial].
Details of this approach, including how we used Kronecker products over Gaussian Markov random fields to define appropriate random effects, are provided in Appendix A.1.2.
The best performing model was selected according to the conditional predictive ordinate (CPO) criteria, a measure of leave-one-out model performance that can be calculated directly in `R-INLA` without explicit model refitting [@held2010posterior].
The \textsc{R} [@r] code used to implement the models and produce  results is available from `github.com/athowes/multi-agyw`.
We used `sf` [@pebesma2018sf] for handling of spatial data, `orderly` [@orderly] for reproducible research, `ggplot2` for data visualisation [@wickham2016ggplot2] and `rticles` [@allaire2022rticles] for reporting via `rmarkdown` [@allaire2022rmarkdown].

## HIV indicators and prevention prioritisation

Using risk group proportion estimates, and subnational estimates developed using the Naomi model, we calculated HIV prevalence, PLHIV, HIV incidence and number of new HIV infections stratified by district, age group and risk group.
Further details are provided in Appendix A.2.

### Prevalence and PLHIV

To calculate HIV prevalence by risk group, we disaggregated the district-age specific prevalence estimates from Naomi estimates to risk groups using odds ratios for the relative odds of having HIV between risk groups, calculated from a logistic regression of country-age specific household survey HIV bio-marker data.
<!-- TODO: Be more specific about the data source here. -->
Prevalence disaggregation was on the logit scale to ensure that HIV prevalence in each risk group remained in the range 0\% to 100\%.
The number of PLHIV was calculated by multiplying HIV prevalence by risk group population size.

### Incidence and new infections

We disaggregated HIV incidence by risk group using the HIV infection risk ratios in Table \ref{tab:risk-groups}.
The risk ratio used for the not sexually active risk group was zero, excluding incidence from non-sexual transmission which is negligible in these populations.
For the non-regular or multiple partner(s) risk group, the risk ratio was based on analyses of risk factors for incident HIV infection from studies in sub-Saharan Africa [@slaymaker2020croi] and supported by a recent systematic review [@jia2022risk].
Risk ratios for the highest risk group vary based upon general population HIV incidence and are based on an analysis of HIV prevalence among FSW relative to population prevalence [@aidsinfo] using data from the UNAIDS Key Population Atlas [@kpatlas].
The number of new HIV infections were calculated by multiplying HIV incidence by susceptible risk group population size.

### Infections reached

For each possible stratification of risk, we calculated the expected number of new infections that would be found per person reached when prioritising according to incidence.
To do so, we ordered the strata by descending incidence before cumulatively summing the expected new infections and population.
We assumed it was possible to reach all members of every strata.

# Results

<!-- Suggestions for other things to include in the results from Katie: -->
<!-- * Number in different buckets of intervention strategies (caveat on not including STI past year) -->
<!-- * Section on incidence disaggregated by risk group -->

## Data

```{r fig.cap="Surveys used in our analysis by year, survey type, sample size, and whether the survey included a question about transactional sex. \\label{fig:available-surveys}"}
knitr::include_graphics("depends/available-surveys.jpeg")
```

```{r results='hide', warning=FALSE, message=FALSE}
available_surveys <- read_csv("depends/available-surveys.csv")

#' What was the total number of surveys?
total_surveys <- available_surveys %>%
  pull(survey_id) %>%
  unique() %>%
  length()

#' How many AGYW, in total, were surveyed?
agyw_surveyed <- available_surveys %>%
  pull(Y015_029) %>%
  sum()

#' What about disaggregating these AGYW by age?
agyw_surveyed_age <- available_surveys %>%
  summarise(
    Y015_019 = sum(Y015_019),
    Y020_024 = sum(Y020_024),
    Y025_029 = sum(Y025_029)
  )

#' What was the total number of surveys including a transactional question?
giftsvar_surveys <- available_surveys %>%
  filter(giftsvar == 1) %>%
  pull(survey_id)

length(giftsvar_surveys)

#' How many AGYW were sampled in these surveys?
agyw_surveyed_giftsvar <- available_surveys %>%
  filter(survey_id %in% giftsvar_surveys) %>%
  pull(Y015_029) %>%
  sum()

#' Disaggregated by age
#' This isn't quite right because some 25-29 are not asked FSW question in DHS surveys, which I fix below
agyw_surveyed_giftsvar_age <- available_surveys %>%
  filter(survey_id %in% giftsvar_surveys) %>%
  summarise(
    Y015_019 = sum(Y015_019),
    Y020_024 = sum(Y020_024),
    Y025_029 = sum(Y025_029)
  )

#' These surveys excluded AGYW 25-29 for the transactional sex question
#' I've quickly hard-coded this by looking at data-check.pdf from fit_fsw-logit-sae
#' In future it could be automatically updated based on that report instead
giftsvar_exclude_Y025_029 <- c("CMR2018DHS", "MWI2015DHS", "UGA2016DHS", "ZAF2016DHS", "ZMB2018DHS", "ZWE2015DHS")

giftsvar_Y025_029_overcount <- available_surveys %>%
  filter(
    survey_id %in% giftsvar_exclude_Y025_029,
  ) %>%
  pull(Y025_029) %>%
  sum()

#' What was the median number of surveys by country, and the range?
number_of_surveys <- available_surveys %>%
  group_by(iso3) %>%
  summarise(
    count = length(unique(survey_id))
  )

number_of_surveys %>%
  summarise(
    min = min(count),
    median = median(count),
    max = max(count)
  )
```

We included `r total_surveys` surveys in our analysis (Fig \ref{fig:available-surveys}, Table B.3), with a total sample size of `r comma(agyw_surveyed)` women aged 15-29 years (`r comma(agyw_surveyed_age$Y015_019)` aged 15-19 years, `r comma(agyw_surveyed_age$Y020_024)` aged 20-24 years, and `r comma(agyw_surveyed_age$Y025_029)` aged 25-29 years).
Of these, `r length(giftsvar_surveys)` surveys included a specific transactional sex question, with a total sample size of `r comma(agyw_surveyed_giftsvar - giftsvar_Y025_029_overcount)` (`r comma(agyw_surveyed_giftsvar_age$Y015_019)` aged 15-19 years, `r comma(agyw_surveyed_giftsvar_age$Y020_024)` aged 20-24 years, and `r comma(agyw_surveyed_giftsvar_age$Y025_029 - giftsvar_Y025_029_overcount)` aged 25-29 years^[There were `r length(giftsvar_exclude_Y025_029)` DHS surveys which excluded women 25-29 from the transactional sex survey question.]).
The median number of surveys per country was four, ranging from one in Botswana and South Africa to six in Uganda.

## Model selection and model fit

The best fitting multinomial regression model included correlated spatial effects and first-order autoregressive temporal effects, but not spatio-temporal interactions.
The best logistic regression model for transactional sex included correlated spatial effects and the proportion of men who reported ever paying for sex covariate.
Model performance according to the CPO criteria is provided in Appendix A.1.2.4 for all considered models.
Direct estimates of the four risk group proportions from these surveys were highly correlated with our modelled estimates at a national-level.

## Risk group estimates

```{r continental-map, fig.cap="The spatial distribution (posterior mean) of the AGYW risk group proportions in 2018. Countries in grey were not included in our analysis. \\label{fig:continental-map}"}
#' Note that the subnational borders look a bit overpowering when the quality of this is reduced
knitr::include_graphics("depends/3p1-continental-map.jpeg")
```

```{r within-between-country-variation, fig.cap="National (in white) and subnational (in color) posterior means of the risk group proportions. We used regions of sub-Saharan Africa defined according to the UN geocode (M49). \\label{fig:within-between-country-variation}"}
knitr::include_graphics("depends/3p1-within-between-country-variation.jpeg")
```

Fig \ref{fig:continental-map} and Fig \ref{fig:within-between-country-variation} show posterior mean estimates for the proportion in each risk group from our final model.
Posterior mean, median, and 95\% credible intervals are provided in the supplementary material.
In subsequent results, all estimates refer to 2018, the most recent year included in our analysis, unless otherwise indicated.

```{r results=FALSE}
df_3p1_subnational <- read_csv("depends/df_3p1_subnational.csv")
df_3p1_national <- read_csv("depends/df_3p1_national.csv")

#' Quantification of points discussed
#' District level quantile information (age disaggregated)
district_age_quantiles <- df_3p1_subnational %>%
  select(iso3, indicator, age_group, estimate_smoothed) %>%
  group_by(indicator, age_group) %>%
  summarise(
    lower = 100 * quantile(estimate_smoothed, probs = 0.025, na.rm = TRUE),
    median = 100 * median(estimate_smoothed, na.rm = TRUE),
    upper = 100 * quantile(estimate_smoothed, probs = 0.975, na.rm = TRUE)
  )

district_Y015_019_nosex12m_quantiles <- filter(district_age_quantiles, indicator == "No sex", age_group == "15-19")

#' District level quantile information aggregated by age (population weighted)
district_quantiles <- df_3p1_subnational %>%
  select(iso3, area_id, indicator, age_group, estimate_smoothed, population_mean) %>%
  group_by(area_id, indicator) %>%
  summarise(
    estimate_smoothed = sum(estimate_smoothed * population_mean) / sum(population_mean)
  ) %>%
  ungroup() %>%
  group_by(indicator) %>%
  summarise(
    lower = 100 * quantile(estimate_smoothed, probs = 0.025, na.rm = TRUE),
    median = 100 * median(estimate_smoothed, na.rm = TRUE),
    upper = 100 * quantile(estimate_smoothed, probs = 0.975, na.rm = TRUE)
  )

#' What proportion of 15-19 year-olds are cohabiting (or other indicators) in MOZ?
moz_Y015_019 <- df_3p1_national %>%
  filter(
    age_group == "15-19",
    iso3 == "Mozambique"
  ) %>%
  mutate(estimate_smoothed = 100 * estimate_smoothed)

moz_Y015_019_nosex12m <- filter(moz_Y015_019, indicator == "No sex")
moz_Y015_019_sexcohab <- filter(moz_Y015_019, indicator == "One cohabiting partner")
moz_Y015_019_sexnonreg <- filter(moz_Y015_019, indicator == "Non-regular or multiple partner(s)")

#' In which countries are the majority of 15-19 year-olds not sexually active?
#' (Just MOZ!)
majority_claim_moz <- df_3p1_national %>%
  filter(
    age_group == "15-19",
    indicator == "No sex"
  ) %>%
  mutate(estimate_smoothed = 100 * estimate_smoothed) %>%
  filter(estimate_smoothed < 50)

#' What proportion of 20-29 year-olds are cohabiting versus with nonregular partner(s),
#' above and below the south / east dividing border (UN geoscheme)?
dividing_border <- df_3p1_subnational %>%
  mutate(
    border = ifelse(region %in% c("Eastern", "Central"), "Above", "Below")
  ) %>%
  select(iso3, area_id, indicator, age_group, border, estimate_smoothed, population_mean) %>%
  filter(
    indicator %in% c("One cohabiting partner", "Non-regular or multiple partner(s)"),
    age_group %in% c("20-24", "25-29")
  ) %>%
  group_by(area_id, indicator) %>%
  summarise(
    estimate_smoothed = sum(estimate_smoothed * population_mean) / sum(population_mean),
    border = max(border) #' Should just be a way to keep border column
  ) %>%
  ungroup() %>%
  group_by(border, indicator) %>%
  summarise(
    lower = 100 * quantile(estimate_smoothed, probs = 0.025, na.rm = TRUE),
    median = 100 * median(estimate_smoothed, na.rm = TRUE),
    upper = 100 * quantile(estimate_smoothed, probs = 0.975, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(across(lower:upper, ~round(.x, digits = 1)))

dividing_border_above_sexcohab <- filter(dividing_border, border == "Above", indicator == "One cohabiting partner")
dividing_border_above_sexnonreg <- filter(dividing_border, border == "Above", indicator == "Non-regular or multiple partner(s)")
dividing_border_below_sexcohab <- filter(dividing_border, border == "Below", indicator == "One cohabiting partner")
dividing_border_below_sexnonreg <- filter(dividing_border, border == "Below", indicator == "Non-regular or multiple partner(s)")

#' What's the average FSW proportion by age?
fsw_age <- df_3p1_subnational %>%
  filter(indicator == "FSW") %>%
  group_by(iso3, age_group) %>%
  summarise(estimate_smoothed = sum(estimate_smoothed * population_mean, na.rm = TRUE) / sum(population_mean, na.rm = TRUE)) %>%
  group_by(age_group) %>%
  summarise(
    lower = 100 * quantile(estimate_smoothed, probs = 0.025, na.rm = TRUE),
    median = 100 * median(estimate_smoothed, na.rm = TRUE),
    upper = 100 * quantile(estimate_smoothed, probs = 0.975, na.rm = TRUE)
  ) %>%
  mutate(across(lower:upper, ~ round(.x, digits = 1)))

fsw_age_Y015_019 <- filter(fsw_age, age_group == "15-19")
fsw_age_Y020_024 <- filter(fsw_age, age_group == "20-24")
fsw_age_Y025_029 <- filter(fsw_age, age_group == "25-29")

#' What about by country?
fsw_iso3 <- df_3p1_subnational %>%
  filter(indicator == "FSW") %>%
  group_by(iso3) %>%
  summarise(estimate_smoothed = sum(estimate_smoothed * population_mean, na.rm = TRUE) / sum(population_mean, na.rm = TRUE))
```

The median national FSW proportion was `r fsw_age_Y015_019$median`\% (95\% CI `r fsw_age_Y015_019$lower`--`r fsw_age_Y015_019$upper`) for the 15-19 age group, `r fsw_age_Y020_024$median`\% (95\% CI `r fsw_age_Y020_024$lower`--`r fsw_age_Y020_024$upper`) for the 20-24 age group and `r fsw_age_Y025_029$median`\% (95\% CI `r fsw_age_Y025_029$lower`--`r fsw_age_Y025_029$upper`) for the 25-29 age group.

<!-- I would like some more quantitive results here comparing the amount of within-country variation vs. between country variation. -->

In the 20-24 and 25-29 year age groups, the majority of women were either cohabiting or had non-regular or multiple partner(s).
Countries in eastern and central Africa (Cameroon, Kenya, Malawi, Mozambique, Tanzania, Uganda, Zambia and Zimbabwe) had a higher proportion of women in these age groups cohabiting (`r dividing_border_above_sexcohab$median`\% [95\% CI `r dividing_border_above_sexcohab$lower`--`r dividing_border_above_sexcohab$upper`\%] as compared with `r dividing_border_above_sexnonreg$median`\% [95\% CI `r dividing_border_above_sexnonreg$lower`--`r dividing_border_above_sexnonreg$upper`\%] with non-regular partner[s]).
In contrast, countries in southern Africa (Botswana, Eswatini, Lesotho, Namibia and South Africa) had a higher proportion with non-regular or multiple partner(s) (`r dividing_border_below_sexnonreg$median`\% [95\% CI `r dividing_border_below_sexnonreg$lower`--`r dividing_border_below_sexnonreg$upper`\%], compared with `r dividing_border_below_sexcohab$median`\% [95\% CI `r dividing_border_below_sexcohab$lower`--`r dividing_border_below_sexcohab$upper`\%] cohabiting).
This clear geographic delineation passes along the border of Mozambique, through the interior of Zimbabwe and along the border of Zambia (Fig \ref{fig:continental-map}).

In most districts (`r round(district_Y015_019_nosex12m_quantiles$median, 1)`\%; 95\% credible interval [CI] `r round(district_Y015_019_nosex12m_quantiles$lower, 1)`--`r round(district_Y015_019_nosex12m_quantiles$upper, 1)`) adolescent girls aged 15-19 were not sexually active.
The exception was Mozambique, where the majority (`r 100 - round(moz_Y015_019_nosex12m$estimate_smoothed, 2)`\%) were sexually active in the past year and close to a third (`r round(moz_Y015_019_sexcohab$estimate_smoothed, 2)`\%) were cohabiting with a partner.

```{r}
variance_proportions <- read_csv("depends/variance-proportions-uncertainty.csv")
variance_proportions_age <- 100 * filter(variance_proportions, variable == "age") %>% select(-variable) %>% round(digits = 3)
variance_proportions_iso3 <- 100 * filter(variance_proportions, variable == "iso3") %>% select(-variable) %>% round(digits = 3)
variance_proportions_area <- 100 * filter(variance_proportions, variable == "area") %>% select(-variable) %>% round(digits = 3)
variance_proportions_year <- 100 * filter(variance_proportions, variable == "year") %>% select(-variable) %>% round(digits = 3)
variance_proportions_cat <- 100 * filter(variance_proportions, variable == "cat") %>% select(-variable) %>% round(digits = 3)
```

Age group was the most important factor explaining variation in risk group proportions, and accounted for `r variance_proportions_age$mean`\% (95\% CI `r variance_proportions_age$lower`--`r variance_proportions_age$upper`) of total variation.
The primary change in risk group proportions by age group occurs between the 15-19 age group and 20-29 age group (Fig \ref{fig:within-between-country-variation}).
The next most important factor was location.
Country-level differences explained `r variance_proportions_iso3$mean`\% (95\% CI `r variance_proportions_iso3$lower`--`r variance_proportions_iso3$upper`) of variation, while district-level variation within countries explained `r variance_proportions_area$mean`\% (95\% CI `r variance_proportions_area$lower`--`r variance_proportions_area$upper`).
<!-- Can I plot the country-level and district-level random effects to visualise this? -->
Temporal changes only explained `r variance_proportions_year$mean`\% (95\% CI `r variance_proportions_year$lower`--`r variance_proportions_year$upper`) of variation, indicating very little change in risk group proportions over time.
We observed similar variance decomposition results fitting each country individually, and using other model specifications.

<!-- Are these results presented? (A reviewer might ask for it if referencing here; but fine to leave as this statement for now.) -->

<!-- TODO: Write another paragraph here calling out some examples of the amount of variation -->
<!-- Awaiting additional results: https://github.com/athowes/multi-agyw/issues/95 -->

## Expected infections reached

```{r fig.cap="Percentage of new infections reached across all 13 countries, taking a variety of risk stratification approaches, against the percentage of at risk population required to be reached. \\label{fig:infections-reached}"}
knitr::include_graphics("depends/infections-reached.jpeg")
```

```{r results=FALSE}
inf <- read_csv("depends/infections-reached.csv")
inf_country <- readRDS("depends/infections-reached-country.rds")

#' What is the minimum proportion of the population to reach in order to
#' achieve reaching over q \in [0, 1] proportion of the new infections?
min_pop_reached <- function(inf, q) {
  inf %>%
    filter(prop_infections_averted_cumulative > q) %>%
    group_by(stratification) %>%
    filter(prop_population_cumulative == min(prop_population_cumulative)) %>%
    select(stratification, prop_infections_averted_cumulative, prop_population_cumulative)
}

inf_quarter <- min_pop_reached(inf, 0.25)
inf_half <- min_pop_reached(inf, 0.5)

inf_half_with_behaviour <- round(100 * filter(inf_half, stratification == "Area, age, behaviour")$prop_population_cumulative, 1)
inf_half_without_behaviour <- round(100 * filter(inf_half, stratification == "Area, age")$prop_population_cumulative, 1)

#' At a national level
country_min_pop <- lapply(inf_country, function(x) min_pop_reached(x, q = 0.25)) %>%
  bind_rows(.id = "country") %>%
  filter(stratification %in% c("Area, age, behaviour", "Area, age"))

inf_country_half <- country_min_pop %>%
  group_by(stratification) %>%
  summarise(
    max = round(100 * max(prop_population_cumulative), 1),
    median = round(100 * median(prop_population_cumulative), 1),
    min = round(100 * min(prop_population_cumulative), 1)
  )

inf_country_half_with_behaviour <- filter(inf_country_half, stratification == "Area, age, behaviour")
inf_country_half_without_behaviour <- filter(inf_country_half, stratification == "Area, age")

#' What proportion of new infections are among FSW as compared with the proportion the population that are FSW?
inf_fsw <- inf %>%
  filter(
    stratification == "Behaviour",
    category == "sexpaid12m"
  ) %>%
  select(prop_infections_averted_cumulative, prop_population_cumulative) %>%
  mutate(across(everything(), ~ round(100 * .x, digits = 1)))

#' What about by country?
inf_fsw_country <- lapply(inf_country, function(x)
  x %>%
    filter(stratification == "Behaviour", category == "sexpaid12m") %>%
    select(prop_infections_averted_cumulative, prop_population_cumulative) %>%
      mutate(across(prop_infections_averted_cumulative:prop_population_cumulative, ~ 100 * round(.x, 3)))
) %>%
  bind_rows(.id = "country")

inf_fsw_kenya <- filter(inf_fsw_country, country == "Kenya")
inf_fsw_tanzania <- filter(inf_fsw_country, country == "Tanzania")
```

For any given fraction of AGYW prioritised, substantially more new infections were reached by strategies that included behavioural risk stratification.
Reaching half of all expected new infections required reaching `r inf_half_without_behaviour`\% of the population when stratifying by subnational area and age, but only `r inf_half_with_behaviour`\% when behavioural stratification was included (Fig \ref{fig:infections-reached}).
The majority of this benefit came from reaching FSW, who were `r inf_fsw$prop_population_cumulative`\% of the population but `r inf_fsw$prop_infections_averted_cumulative`\% of all new infections.

<!-- How does this vary by country / region?  Maybe worth calling out result for West Africa vs. East Africa vs. Southern Africa here to highlight the difference. Also how does this compare with other estimates? And what does Oli think about this? I recognize that comparison is not straightforward because age group difference! -->

Considering each country separately, on average, reaching half of new infections required reaching `r inf_country_half_without_behaviour$median`\% (range `r inf_country_half_without_behaviour$min`-`r inf_country_half_without_behaviour$max`\%) of the population when stratifying by area and age, reducing to `r inf_country_half_with_behaviour$median`\% (range `r inf_country_half_with_behaviour$min`-`r inf_country_half_with_behaviour$max`\%) when behaviour was included.
The relative importance of stratifying by age, location and behaviour varied between countries, analogous to the varying contribution of each to the total variance (Appendix B).
For example, FSW in Kenya were estimated to be `r inf_fsw_kenya$prop_population_cumulative`\% of the population and close to a third (`r inf_fsw_kenya$prop_infections_averted_cumulative`\%) of all new infections, whereas FSW in Tanzania were just `r inf_fsw_tanzania$prop_population_cumulative`\% of the population and `r inf_fsw_tanzania$prop_infections_averted_cumulative`\% of all new infections.

<!-- Are we able to add an analysis of the number who are prioritized for the 'standard' and 'enhanced' prevention packages in the Global Strategy? That gets a lot of emphasis in the Introduction, so would be nice to report about here. -->

---
abstract: |
  **Background:**
  The Global AIDS Strategy 2021-2026 identifies adolescent girls and young women (AGYW) as a priority population for HIV prevention, and recommends differentiated intervention portfolios geographically based on local HIV incidence and individual risk behaviours.
  We aimed to estimate prevalence of HIV risk behaviours and associated HIV incidence at a small spatial scale among AGYW living in 13 countries in sub-Saharan Africa.

  **Methods:**
  We analysed `r total_surveys` geospatially referenced national household surveys conducted between 1999-2018 across 13 high burden countries in sub-Saharan Africa.
  Female survey respondents aged 15-29 years were classified into four risk groups (not sexually active, cohabiting, non-regular or multiple partner[s] and female sex workers [FSW]) based on reported sexual behaviour.
  We used a Bayesian spatio-temporal multinomial regression model to estimate the proportion of AGYW in each risk group stratified by district, year, and five-year age group.
  Using subnational estimates of HIV prevalence and incidence produced by countries with support from UNAIDS, we estimated new HIV infections in each risk group by district and age group.
  We then assessed the efficiency of prioritising interventions according to risk group.

  **Results:**
  Data consisted of `r comma(agyw_surveyed)` female survey respondents aged 15-29.
  Among women aged 20-29, cohabiting (`r dividing_border_above_sexcohab$median`\%) was more common in eastern Africa than non-regular or multiple partner(s) (`r dividing_border_above_sexnonreg$median`\%), while in southern countries non-regular or multiple partner(s) (`r dividing_border_below_sexnonreg$median`\%) were more common than cohabiting (`r dividing_border_below_sexcohab$median`\%).
    Risk group proportions varied substantially across age groups (`r variance_proportions_age$mean`\% of total variation explained), countries (`r variance_proportions_iso3$mean`\%), and between districts within each country (`r variance_proportions_area$mean`\%), but changed little over time (`r variance_proportions_year$mean`\%).
  Prioritisation based on behavioural risk, in combination with location- and age-based prioritisation, reduced the proportion of population required to be reached in order to find half of all expected new infections from `r inf_half_without_behaviour`\% to `r inf_half_with_behaviour`\%.
  FSW were `r inf_fsw$prop_population_cumulative`\% of the population but `r inf_fsw$prop_infections_averted_cumulative`\% of all expected new infections.

  **Conclusions:**
  Our risk group estimates provide data for HIV programmes to set targets and implement differentiated prevention strategies outlined in the Global AIDS Strategy.
  Successfully implementing this approach would result in more efficiently reaching substantially more of those at risk for infections.

author_summary: |
  Exposure to HIV among adolescent girls and young women depends on the level of unsuppresed HIV in their local community among potential partners, and their individual sexual behaviour.
  To support HIV programmes in identifying and providing services to those at highest risk, we geographically mapped HIV risk groups using sexual risk behaviour data from household surveys and estimated the proportion of people living with HIV in each risk group, and the rate of infections in each risk group.
  Doing so also allowed us to estimate the the proportion of people living with HIV in each risk group, as well as the rate of new infections in each risk group.
  We found that prioritising HIV prevention using behavioural risk groups almost halved the proportion of the population that would need to be reached to find half of all expected new infections.
---

# Discussion

<!-- Suggestions for other things to include in the discussion from Katie: -->
<!-- *  Discussion of different interventions to be implemented in the different risk groups based on Global Strategy -->

We estimated the proportion of AGYW who fall into different risk groups at a district level in 13 sub-Saharan African countries.
Our estimates support consideration of differentiated prevention programming according to geographic locations and risk behaviour, as outlined in the Global AIDS Strategy.
<!-- Some type of comment about how novel this would be if implemented -->

We defined four risk groups based on sexual behaviour, the most proximal determinant of risk.
Other factors, such as condom usage or type of sexual act, may account for additional heterogeneity in risk from sexual behaviour.
However, we did not include these factors in view of measurement difficulties, consistency across contexts, and the operational benefits of describing risk parsimoniously.
<!-- Reasons why we didn't include condom usage: -->
<!-- Defining condom usage is challenging, particularly for those with multiple partners, and reporting is likely biased due to social desirability. -->
<!-- Additionally, consistent condom usage is rare amongst those cohabiting, and as such likely only relevant for those with non-regular sexual partners.  -->
Sexual behaviour confers risk only when AGYW reside in geographic locations where there is unsuppressed viral load among their potential partners.

We did not include more distal determinants, such as school attendance, orphanhood, or gender empowerment, as we expect their effects on risk to largely be mediated by more proximal determinants.
However, to effectively implement programming, it is crucial to understand these factors, as well as the broader structural barriers and limits to personal agency faced by AGYW.
Importantly, programs must ensure that intervention prioritisation occurs without stigmatising or blaming AGYW.

Brugh et al. [@brugh2021characterizing] previously geographically mapped AGYW HIV risk groups using biomarker and behavioural data from the most recent surveys in Eswatini, Haiti and Mozambique to define and subsequently map risk groups with a range of machine learning techniques.
Our work builds on Brugh et al. [@brugh2021characterizing] by including more countries, integrating a greater number of surveys, and connecting risk group proportions with HIV epidemic indicators to help inform programming.

By considering a range of possible risk stratification strategies, we showed that successful implementation of a risk-stratified approach would allow substantially more of those at risk for infections to be identified before infection occurs.
A considerable proportion of estimated new infections were among FSW, supporting the case for HIV programming efforts focused on key population groups [@baral2012burden].
There is substantial variation in the importance of prioritisation by age, location and behaviour within each country.
This highlights the importance of understanding and tailoring HIV prevention efforts to country-specific contexts.
By standardising our analysis across all 13 countries, we showed the additional efficiency benefits of resource allocation between countries.

We found a geographic delineation in the proportion of women cohabiting between southern and eastern Africa, calling attention to a divide attributable to many cultural, social and economic factors.
The delineation does not represent a boundary between predominately Christian and Muslim populations, which is further north.
We also note that the high numbers of adolescent girls aged 15-19 cohabiting in Mozambique is markedly different from the other countries [@unicef].
<!-- Source suggested by KJ on this: https://onlinelibrary.wiley.com/doi/10.1111/padr.12043 -->
<!-- Widespread changes in sexual behaviour in eastern and southern Africa: Challenges to achieving global HIV targets? Longitudinal analyses of nationally representative surveys -->

Our modelled estimates of risk group proportions improve upon direct survey results for three reasons.
First, by taking a modular modelling approach, we integrated all relevant survey information from multiple years, allowing estimation of the FSW proportion for surveys without a specific transactional sex question.
Second, whereas direct estimates exhibit large sampling variability at a district level, we alleviated this issue using spatio-temporal smoothing (Fig B.2).
Third, we provided estimates in all district-years, including those not directly sampled by surveys, allowing estimates to be consistently fed into further analysis and planning pipelines (such as our analysis of risk group specific prevalence and incidence).

The final surveys included in our risk model model were conducted in 2018.
We plan to update our analysis with more surveys as they become available, but do not anticipate that the risk group proportions will change substantially, as we found that they did not change significantly over time.

Our analysis focused on females aged 15-29 years, and could be extended to consider optimisation of prevention more broadly, accounting for the `r 100 - infections_percent`\% of new infections among adults 15-49 which occur in women 30-49 and men 15-49.
Estimating sexual risk behaviour in adults 15-49 would be a crucial step toward greater understanding of the dynamics of the HIV epidemic in sub-Saharan Africa, and would allow incidence models to include stratification of individuals by sexual risk.

<!-- Paragraph about AGFS e.g. our results are consistent with age at first sex findings [@nguyen2022trends]. -->

## Limitations

Our analysis was subject to challenges shared by most approaches to monitoring sexual behaviour in the general population [@cleland2004monitoring].
In particular, under-reporting of higher risk sexual behaviours among AGYW could affect the validity of our risk group proportion estimates.
Due to social stigma or disapproval, respondents may be reluctant to report non-marital partners [@nnko2004secretive, @helleringer2011reliability] or may bias their reporting of sexual debut [@zaba2004age; @wringe2009comparative; @nguyen2022trends].
For guidance of resource allocation, differing rates of under-reporting by country, district, year or age group are particularly concerning to the applicability of our results; and, while it may be reasonable to assume a constant rate over space-time, the same cannot be said for age, where aspects of under-reporting have been shown to decline as respondents age [@glynn2011assessing], suggesting that the elevated risks we found faced by younger women are likely a conservative estimate.
If present, these reporting biases will also have distorted the estimates of infection risk ratios and prevalence ratios we used in our analysis, likely over-attributing risk to higher risk groups.

We have the least confidence in our estimates for the FSW risk group.
As well as having the smallest sample sizes, our transactional sex estimates do not overcome the difficulties of sampling hard to reach groups.
In adjusting our estimates of transactional sex to match national FSW estimates, we inherent their limitations. 
Furthermore, we do not consider seasonal migration patterns, which may particularly effect FSW size.

The efficiency of each stratified prevention strategy depends on the ability of programmes to identify and effectively reach those in each strata.
Our analysis of new infections potentially averted assumed a "best-case" scenario where AGYW of every strata can be reached perfectly, and should therefore be interpreted as illustrating the potentially obtainable benefits rather than benefits which would be obtained from any specific intervention strategy.
In practice, stratified prevention strategies are likely to be substantially less efficient than this best-case scenario.
Factors we did not consider include the greater administrative burden of more complex strategies, variation in difficulty or feasibility of reaching individuals in each strata, variation in the range or effectiveness of interventions by strata, and changes in strata membership that may occur during the course of a year.
Identifying and reaching behavioural strata may be particularly challenging.
Empirical evaluations of behavioural risk screening tools have found only moderate discriminatory ability [@jia2022risk], and risk behaviour may change rapidly among young populations, increasing the challenge to effectively deliver appropriately timed prevention packages.
This consideration may motivate selecting risk groups based on easily observable attributes, such as attendance of a particular service or facility, rather than sexual behaviour.

## Conclusion

We estimated the proportion of AGYW aged 15-19, 20-24 and 25-29 years in four sexual risk categories at a district-level in 13 priority countries.
Using these estimates, we performed an analysis of the number of infections that could be reached by prioritisation based upon location, age and behaviour.
Though subject to limitations, our results quantify the crucial importance of prioritising reaching AGYW in the highest risk districts and at highest behavioural risk.
<!-- TODO: Add "... and highlight the need to consider country-specific contexts." once there is more discussion of this in main text. -->

## Supporting information

**Appendix A**: Statistical modelling \newline
**Appendix B**: Supplementary tables and figures

## Funding

<!-- A first version of this work was funded by the Global Fund (###) via UNAIDS -->

AH was supported by the EPSRC Centre for Doctoral Training in Modern Statistics and Statistical Machine Learning (EP/S023151/1).
AH, VKN, KAR, JWE were supported by the Bill and Melinda Gates Foundation (OPP1190661, OPP1164897).
TMW was supported by Imperial College London (President's PhD Scholarship).
SF was supported by the EPSRC (EP/V002910/2).
JWE was supported by UNAIDS and National Institute of Allergy and Infectious Disease of the National Institutes of Health (R01AI136664). 
This research was supported by the MRC Centre for Global Infectious Disease Analysis (MR/R015600/1), jointly funded by the UK Medical Research Council (MRC) and the UK Foreign, Commonwealth \& Development Office (FCDO), under the MRC/FCDO Concordat program and is also part of the EDCTP2 programme supported by the European Union.

LZ, IW, MM and CB were supported by...

For the purpose of Open Access, the authors have applied a CC BY public copyright licence to any Author Accepted Manuscript (AAM) version arising from this submission.

## Disclaimer

The findings and conclusions in this manuscript are those of the authors and do not necessarily represent the official position of the funding agencies.

# References {#references .unnumbered}