### 2022-09-16 docs/waterloo.Rmd
Abstract for Waterloo

---
title: "Waterloo student conference abstract submission"
author: "Adam Howes, Imperial College London"
date: "16th September 2022"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

**Title**:
A Bayesian multinomial spatio-temporal for HIV risk group proportions.

**Abstract**:
The Global AIDS Strategy 2021-2026 identifies adolescent girls and young women (AGYW) as a priority population for HIV prevention, and recommends differentiating intervention portfolios according to local HIV incidence, and individual risk behaviour.
To help countries implement the strategy, we developed a Bayesian multinomial spatio-temporal model to estimate district-level HIV risk group proportions for AGYW across 13 high burden countries in sub-Saharan Africa.
We used the multinomial-Poisson transformation to enable fast and accruate inference for the model via integrated nested Laplace approximations.
Using our risk group proportions we estimated additional policy-relevant indicators including HIV prevalence and HIV incidence by risk group, as well as numbers of infections that could be averted using different intervention strategies.
For more information, see [`athowes/multi-agyw`](https://github.com/athowes/multi-agyw).