### 2022-09-17 src/docs_04-01-22-update/04-01-22-update.Rmd
update upload

---
title: A multinomial spatio-temporal model for sexual risk behaviour with application to adolescent girls and young women in 13 sub-Saharan African countries
subtitle: HIV Inference Lab Group Meeting
author: Adam Howes
institute: Imperial College London
date: January 2022
bibliography: citations.bib
output:
  beamer_presentation:
    latex_engine: pdflatex
    highlight: haddock
    fig_width: 7 
    fig_height: 3
    includes:
        in_header: preamble.tex
---

##

\begin{table}
\centering
\begin{tabularx}{\textwidth}{lXlX}
\toprule
Category & Description & Risk ratio & Notes \\ 
 \midrule
None & Not sexually active & 0 & Ignores incidence from e.g. drug injection (assumed to be negligible). \\
Low & One cohabiting sexual partner & 1 & Baseline category. \\
High & Non-regular sexual partner(s) & 1.72 & From ALPHA network pooled analysis. Supported by Jia et al. \\
Very high & Young women from key populations & 13 & Sabin et al. \\
\bottomrule
\end{tabularx}
\caption{HIV risk categories and risk ratios.}
\label{tab:risk-categories}
\end{table}

##

```{r echo=FALSE}
knitr::include_graphics("depends/category-flowchart.pdf")
```

##

```{r echo=FALSE}
knitr::include_graphics("depends/available-surveys.pdf")
```

## Statistical model

Fit 13 models of the form:

\begin{align}
    y_{itak}^\star &\sim \text{xPoisson}(\lambda_{itak}) \\
    \log(\lambda_{itak}) &= \beta_k + \alpha_{ak} + \phi_{ik} + \gamma_{tk} + \delta_{itk} + \theta_{ita}.
\end{align}

##

\begin{table}[htb]
\centering
\begin{tabularx}{\textwidth}{llll}
\midrule
Model ID & Spatial structure & Temporal structure & Spatio-temporal interaction \\ 
\midrule
1 & None & None & \xmark \\
2 & IID & None & \xmark \\
3 & Besag & None & \xmark \\
4 & None & IID & \xmark \\
5 & IID & IID & \xmark \\
5x & IID & IID & \cmark \\
6 & Besag & IID & \xmark \\
6x & Besag & IID  & \cmark \\
7 & None & AR1 & \xmark \\
8 & IID & AR1 & \xmark \\
8x & IID & AR1 & \cmark \\
9 & Besag & AR1 & \xmark \\
9x & Besag & AR1 & \cmark \\
\midrule
\end{tabularx}
\end{table}

##

```{r echo=FALSE}
knitr::include_graphics("depends/model-direct-benefits.pdf")
```

## Benefits of smoothing

1. Get estimates even where there isn't direct data

* Some people think of this as "making up data"
* In some sense all estimates are "made up" though
* Uncertainty should be higher in regions with infilling
* Perhaps we do a bad job communicating uncertainty, or it's just hard

## Benefits of smoothing

2. Borrow information across space, surveys, categories

* Space-category effects, survey-category effects, space-survey-category effects (using Kronecker products)
* Fun to think about what models might be reasonable for this and whether structured category effects might be interesting
  * e.g. some model for misclassification of individuals into categories could give a kind of structured category effect (currently it's IID always)
* Wonder to what extent we're oversmoothing
  * e.g. might expect to see discontunities in the data (e.g. urban centers) which we're smoothing away

##

```{r echo=FALSE}
knitr::include_graphics("depends/3-continental-map.png.pdf")
```

##

```{r echo=FALSE}
knitr::include_graphics("depends/aaa-variance-proportions.pdf")
```

##

```{r echo=FALSE}
knitr::include_graphics("depends/age-variation.pdf")
```

## Next steps

* Feedback on paper from people who know more about country contexts
* Connect more to policy. For example by replicating Katie's "number of new infections by risk group" spreadsheet
* Add logistic regression model for proportion of non-regular partner(s) who are KP