### 2022-09-28 src/docs_28-09-22-toronto/28-09-22-toronto.Rmd
Iterate Toronto talk

---
title: Spatio-temporal estimates of HIV risk group proportions for adolescent girls and young women across 13 priority countries in sub-Saharan Africa
subtitle: Centre for Global Health Research, UToronto
author: Adam Howes
institute: Imperial College London
date: September 2022
bibliography: citations.bib
header-includes:
  - \widowpenalties 1 150
output:
  beamer_presentation:
    latex_engine: pdflatex
    highlight: haddock
    fig_width: 7 
    fig_height: 3
    includes:
        in_header: preamble.tex
---

## Background

* In sub-Saharan Africa, adolescent girls and young women (AGYW) aged 15-29 are disproportionately at risk of HIV infection
* This disparity is because of:
  1. Younger age at first sex
  2. Age patterns of sexual mixing
  3. Structural vulnerabilities and power imbalances
  4. Increased susceptibility to HIV infection

## Prevention packages

* Prevention options can be divided into two:
  1. Core package
  2. Intensified interventions
* There are not enough resources to offer the more costly intensified interventions to all AGYW, so it's important to prioritise those at highest risk

## Stratified prevention

::: columns

:::: column

* The Global AIDS strategy 2021-2016 proposed stratifying HIV prevention for AGYW based upon
  1. Population-level HIV incidence
  2. Individual-level sexual risk behaviour
* Takes into account the two most proximal drivers of sexual transmission

::::

:::: column

```{r echo=FALSE, out.width = "125px", fig.cap="Global AIDS strategy"}
knitr::include_graphics("global-aids-strategy.png")
```

::::

:::

## Scope for our work

### Goals

1. Enable implementation of prevention stratified by incidence and behaviour
2. Assess the benefits of such approaches: is it worth it?

### Approach

1. Estimate the proportion of AGYW in four behavioural risk groups at a district level (in 13 countries identified as priority by The Global Fund)
2. Analyze the new infections which could be reached by different stratified prevention strategies

## Data

* We used sexual behaviour data from AIS, BAIS, DHS and PHIA household surveys to place respondents into $K = 4$ risk groups:
  1. $k = 1$ Not sexually active
  2. $k = 2$ One cohabiting sexual partner
  3. $k = 3$ Non-regular sexual partner(s)
  4. $k = 4$ Female sex workers
* District-level HIV incidence, prevalence, population size estimates from the Naomi model [@eaton2021naomi]
* Risk ratios from ALPHA network analysis [@slaymaker2020croi] and UNAIDS analysis led by Keith Sabin

##

```{r echo=FALSE, out.width = "400px"}
knitr::include_graphics("depends/category-flowchart.pdf")
```

##

```{r echo=FALSE, out.width = "400px"}
knitr::include_graphics("depends/available-surveys.png")
```

## Two-stage model for risk group proportions

* Only some of the surveys included a transactional sex question, required to differentiate between the $k = 3$ and $k = 4$ risk groups
* Our approach was to fit a two-stage model
  1. Spatio-temporal multinomial logistic regression model for the proportion of AGYW in the $k = 1, 2, 3^{+}$ risk groups, using all 47 surveys
  2. Spatial logistic regression model for the proportion of those in the $k = 3^{+} = \{3, 4\}$ risk groups who are in the $k = 4$ risk group, using only the 13 surveys with a specific transactional sex question

## Notation

* $k \in \{1, \ldots, 4\}$: risk groups
* $i \in \{1, \ldots, n\}$: districts
* $c[i] \in \{\text{Botswana}, \ldots, \text{Zimbabwe}\}$: country containing district $i$
* $t \in \{1999, \ldots, 2018\}$: years
* $a \in \{\text{15-19}, \text{20-24}, \text{25-29}\}$: age groups
* $y_{ita}^\star = (y_{ita1}^\star, y_{ita2}^\star, y_{ita3}^\star, y_{ita4}^\star)$: survey weighted multinomial observations
* $m_{ita}^\star$: survey weighted multinomial sample size

## Multinomial logistic regression model

* Would like to use integrated nested Laplace approximations for fast, accurate inference, but `R-INLA` is not compatible with multinomial likelihoods because they depend on multiple elements of the latent field
* Instead, use that multinomial logistic regression models can be recast as a Poisson log-linear models using the Poisson trick
* Include observation-specific random effects $\theta_{ita} \sim \mathcal{N}(0, 1000^2)$ in any linear predictor $\eta_{itak} = \theta_{ita} + \cdots$ to ensure exact reproduction of $m_{ita}^\star$

## Multinomial logistic regression model

* Consider models of the form
\begin{align*}
    y_{itak}^\star &\sim \text{xPoisson}(\lambda_{itak}) \\
    \log(\lambda_{itak}) &= \theta_{ita} + \beta_k + \zeta_{c[i]k} + \alpha_{ac[i]k} + \phi_{ik} + \gamma_{tk} + \delta_{itk}.
\end{align*}
* The terms are
  * $\theta_{ita}$: observation (IID)
  * $\beta_k$: category (IID)
  * $\zeta_{ck}$: country-category (IID x IID) 
  * $\alpha_{ack}$: age-country-category (IID x IID x IID)
  * $\phi_{ik}$: space-category ({IID, **Besag**} x IID)
  * $\gamma_{tk}$: year-category ({**IID**, AR1} x IID)
  * $\delta_{itk}$: space-year-category (Implemented but crashing on cluster at the moment...)

## Multinomial logistic regression model

* Independent penalised complexity [@simpson2017penalising] priors on all standard deviation parameters with $\sigma = 0$ and $\mathbb{P}(\sigma > 2.5 = 0.01)$
  * Sidenote, I'm interested as to if joint priors might be more suitable
* Possible (but tricky) to define all these interactions in `R-INLA` by combination of the `group` and `replicate` options
* Used sum-to-zero constraints to make posterior inferences interpretable
  * Because we're interested in the contribution of each random effect to total variance
* Model comparison via CPO statistic

## Logistic regression model

* Consider models of the form
\begin{align*}
    y_{ia4}^\star &\sim \text{Binomial} \left( y_{ia3}^\star + y_{ia4}^\star, q_{ia} \right), \label{eq:logistic-regression} \\
    q_{ia} &= \text{logit}^{-1} \left( \eta_{ia} \right), \\
    \eta_{ia} &= \beta_0 +  \zeta_{c[i]} + \alpha_{ac[i]} + \phi_i + \beta_{\texttt{cfsw}} \texttt{cfsw}_{c[i]}.
\end{align*}
* The terms are
  * $\beta_0$: intercept
  * $\zeta_{c[i]}$: country effects (IID)
  * $\alpha_{ac[i]}$ age-country effects (IID)
  * $\phi_i$: spatial effects (IID, **Besag**)
  * Clients of FSW covariates (`cfswever`, **`cfswrecent`**) [@hodgins2022population]

## Combination and FSW adjustment

* Take 1000 samples from each model, then manipulate suitably to generate estimates for all four risk groups
* We adjusted the samples from the $k = 4$ risk group to match age-country FSW estimates, reallocating into non-regular partner(s)
  * Obtained these by disaggregating @stevens2022estimating by age using estimates of sexually active population from @nguyen2022trends

\begin{center}
\begin{tcolorbox}[width=0.9\textwidth, colframe={title}, colback={white}, title={}]
$\implies$ Estimates of risk group proportions $\rho_{itak}$ by district, year and age group
\end{tcolorbox}
\end{center}

##

```{r echo=FALSE, out.width="350px", fig.cap = "Results of FSW age disaggregation."}
knitr::include_graphics("depends/age-disagg-fsw-line.pdf")
```

##

```{r echo=FALSE, out.width="350px", fig.cap = "PIT histograms and ECDF difference plots."}
knitr::include_graphics("depends/coverage.pdf")
```

##

```{r echo=FALSE, out.width="350px", fig.cap = "We found a geographic discontinuity in behaviour between Southern and Eastern Africa."}
knitr::include_graphics("depends/aids-abstract-A.pdf")
```

##

```{r echo=FALSE, out.width="350px", fig.cap = "Viewing the discontinuity another way."}
knitr::include_graphics("depends/aids-abstract-B.pdf")
```

##

```{r echo=FALSE, out.width="350px", fig.cap = "Proportions of variance explained."}
knitr::include_graphics("depends/aaa-variance-proportions.pdf")
```

## Benefits of our modelled risk group estimates

* Integration of all relevant surveys
  * Two-stage approach allowed estimating FSW proportion even for surveys without a specific transactional sex question
* Alleivating small-sample sizes by borrowing information
  * We borrowed information across space, between countries and over surveys so that our estimates more plausibly reflect spatial heterogeneity
* Estimates where there isn't direct data
  * Although some people think of this as "making up data", the data almost never "speak for themselves" (everything is a model)
  * Uncertainty should be higher in regions with infilling: important to transparently communicate this

##

```{r echo=FALSE, out.width = "350px", fig.cap = "Illustration of the problem with direct survey estimates."}
knitr::include_graphics("depends/model-direct-benefits.png")
```

## HIV prevalence and incidence by risk group

* We used our risk group proportion estimates together with incidence relative risk ratios and prevalence ratios to disaggregate general population HIV estimates in the most recent year
* Disaggregated number of new infections on a linear scale, and people living with HIV (PLHIV) on a logit scale
  * Using a linear scale for PLHIV resulted in prevalences outside $[0, 1]$

\begin{center}
\begin{tcolorbox}[width=0.9\textwidth, colframe={title}, colback={white}, title={}]
$\implies$ Estimates of HIV incidence $\lambda_{iak}$, number of new HIV infections $I_{iak}$, HIV prevalence $\rho_{iak}$ and PLHIV $H_{iak}$ by district, age group and risk group
\end{tcolorbox}
\end{center}

## Prioritisation with risk group information

* Suppose we have all of the information (district, age, and risk group)
* Which are the strata with highest incidence?

```{r include=FALSE}
df <- readr::read_csv("depends/incidence-district-sexbehav.csv")

df_area_age_behav <- df %>%
  select(area_id, age_group, population_nosex12m, incidence_nosex12m, starts_with("population_sex"), starts_with("incidence_sex")) %>%
  pivot_longer(
    cols = starts_with("population") | starts_with("incidence"),
    names_to = c("indicator", "category"),
    names_sep = "_"
  ) %>%
  pivot_wider(
    id_cols = c("area_id", "age_group", "category"),
    names_from = "indicator",
    values_from = "value"
  )
```

```{r echo=FALSE}
df_area_age_behav %>%
  arrange(desc(incidence)) %>%
  head() %>%
  knitr::kable(digits = 2)
```

## Prioritisation without risk group information

* What about if we lost the risk group information? Now what are the strata with the highest incidence?

```{r include=FALSE}
df_area_age <- df_area_age_behav %>%
  group_by(area_id, age_group) %>%
  summarise(
    incidence = sum(incidence * population, na.rm = TRUE) / sum(population, na.rm = TRUE),
    population = sum(population, na.rm = TRUE)
  ) %>%
  ungroup()
```

```{r echo=FALSE}
df_area_age %>%
  select(area_id, age_group, population, incidence) %>%
  arrange(desc(incidence)) %>%
  head() %>%
  knitr::kable(digits = 2)
```

##

```{r echo=FALSE, out.width = "350px", fig.cap = "New infections reached prioritising according to different stratifications."}
knitr::include_graphics("depends/infections-reached.pdf")
```

## Limitations

* Simplistic infections reached analysis
  * No accounting for difficulty in reaching each strata
  * Variable intervention effectiveness
  * Change in strata membership
* Under-reporting of high risk sexual behaviours
  * Variation in under-reporting (likely by age, foremost, and location, less so) particularly concerning
* Risk groups definition justification not clear
  * Didn't consider other important characteristics that may determine risk e.g. condom usage
* Only focused on AGYW 15-29
  * Could be extended to adults of both sexes aged 15-49

## Takeaways

* Risk group estimates can help implement the Global AIDS Strategy; tool and user guide currently being rolled out!
* Importance of reaching FSW
* Countries have different epidemic profiles

## Thanks for listening!

* Joint work with members of the HIV inference group (`hiv-inference.org`) particularly Katie Risher and Jeff Eaton
* The code for this project is at `github.com/athowes/multi-agyw`
* You can find me online at `athowes.github.io`

## Bibliography {.allowframebreaks}