### 2022-12-17 src/docs_review/review.Rmd
Template for (first) response to reviewers

---
title: Response to reviewers for "Spatio-temporal estimates of HIV risk group proportions for adolescent
  girls and young women across 13 priority countries in sub-Saharan Africa"
subtitle: |
  Corresponding author: Adam Howes (`ath19@ic.ac.uk`)
output:
  pdf_document:
    toc: true
    number_sections: true
    keep_tex: true
    includes:
      in_header: preamble.tex
bibliography: citations.bib
---

\newpage

# Reviewer 1 {-}

We thank the reviewer for...

> Comment

Reply, perhaps some code

```{r}
2 + 2
```

\newpage

# Reviewer 2 {-}

We thank the reviewer for...

> Comment

Reply, perhaps some code

```{r}
2 + 2
```

\newpage

# References {#references .unnumbered}