### 2022-12-20 src/docs_review/cover.Rmd
Add cover to editor and update response based on feedback from Katie

---
title: "Cover letter"
author: "Adam Howes"
output:
  pagedown::html_letter:
    self_contained: false
links-to-footnotes: true
paged-footnotes: true
# uncomment this line to produce HTML and PDF in RStudio:
knit: pagedown::chrome_print
---

::: from
Adam Howes    
Email: ath19@ic.ac.uk     
:::

::: date
19th December, 2022
:::

Dear editor,

Thank you for the opportunity to submit a revised version of our manuscript.
We appreciate the time you and the reviewers have given to providing helpful feedback.
Enclosed is a point-by-point response to all points raised.

With kind regards, on behalf of all the co-authors,

Adam Howes