### 2023-01-16 src/docs_review/cover.Rmd

---
title: "Cover letter"
author: "Adam Howes"
output:
  pagedown::html_letter:
    self_contained: false
links-to-footnotes: true
paged-footnotes: true
# uncomment this line to produce HTML and PDF in RStudio:
knit: pagedown::chrome_print
---

::: from
Adam Howes    
Email: ath19@ic.ac.uk     
:::

::: date
19th December, 2022
:::

Dear editor,

Many thanks for your consideration of our manuscript "Spatio-temporal estimates of HIV risk group proportions for adolescent girls and young women across 13 priority countries in sub-Saharan Africa".
We appreciate the thoughtful comments from the reviewers and the opportunity to submit a revised version of our manuscript which incorperates the feedback.
Please find enclosed a point-by-point response to each of the reviewer's comments, and the editorial comments on journal requirements.

With kind regards, on behalf of all the co-authors,

Adam Howes