### 2023-01-16 src/docs_review/review.Rmd
More edits to review response

---
title: Response to reviewers and editorial comments for "Spatio-temporal estimates of HIV risk group proportions for adolescent
  girls and young women across 13 priority countries in sub-Saharan Africa"
subtitle: |
  Corresponding author: Adam Howes (`ath19@ic.ac.uk`)
output:
  pdf_document:
    toc: true
    number_sections: true
    keep_tex: true
    includes:
      in_header: preamble.tex
bibliography: citations.bib
---

\newpage

# Reviewer 1 {-}

We thank the reviewer for their helpful comments regarding the statistical modelling.

> *1. In the manuscript, explain why you use the INLA not WINBUGS coding? The multinomial regression could be modeled in WINBUGS directly.*

The reviewer is right to note that multinomial logistic regression models can be implemented in probabilistic programming languages like `WinBUGS` directly.
However, for this application, Markov chain Monte Carlo approaches would be prohibitively expensive.
For this reason, we chose to use integrated nested Laplace approximations via `R-INLA`, which have been shown to have comparable accuracy for latent Gaussian models in the realistic, pre-asymptotic regime.
We have added the following text to the methods section of the manuscript to clarify this point:

> For models with a Gaussian latent field, INLA has comparable accuracy to Markov chain Monte Carlo in the realistic, pre-asymptotic regime, and is substantially more computationally tractable for high dimensional models like ours, which has 940 districts, 20 years, 3 age groups, and 4 risk groups.

> *2. Is there no other potential covariate that could be used for better modeling?*

Many of the covariates which one might expect to be most predictive of risk group proportions are themselves difficult to accurately measure, and can only lead to modest at best improvements in model performance.
For example, despite the case for their being a clear link between the "proportion clients of FSW" covariate and the "proportion FSW" outcome we found only marginal benefits to inclusion.

We agree that identifying predictive and measureable predictors of risk group proportions, or high risk locations, is an important area for further research.
We have commented on this in the discussion section:

> Comment here.

> *3. Please explain sub-national effect more clearly. Why and how you used it?*

Using district-level spatial random effects allowed us to account for subnational variation in risk group proportions.
We considered both spatially unstructured (IID) and spatially structured (Besag) random effects, which can be implemented in `R-INLA` by setting `model = "iid"` or `model = "besag"` respectively.

> *4. Why the interaction term for spatiotemporal effect didn't consider in the modeling framework?*

```{r}
#' TODO Calculate the average Sobol index for survey random effects
sobol <- 1
```

Fitting each country individually, the proportion of variance (Sobol' index) attributable to the temporal (survey) random effects was on average `r sobol`/% (Supplementary Figure B.3).
This is corroborated by the lack of temporal trends in Figures B.5 through B.17 which show the modelled and direct estimates for each country individually, as well as the fact that unstructured (IID) rather than structured (AR1) temporal random effects were preferred in the model selection (Supplementary Figure A.1).

We had wanted to include the spatiotemporal interaction terms regardless, first for statistical completeness, and second to demonstrate how complex three-way^[As all random effects in the model are interacted with category.] interactions can be specified in `R-INLA` by careful use of the `group` and `replicate` options.
However, we found that including these terms substantially increased the computational burden of the model when fitting all countries jointly, so in the end decided to exclude them.

\newpage

# Reviewer 2 {-}

> *This is a well-crafted manuscript investigating the spatio-temporal estimates of HIV risk group proportions for adolescent girls and young women across 13 priority countries in sub-Saharan Africa. Their analyses identify specific age groups at the district level that should be targeted for HIV intervention in SSA. This is critical in reducing the HIV epidemic in the southern region of SSA. In addition to the main figures, the supplementary Tables and Figures show country-by-country risk, mostly among female sex workers for all age groups. With the help of their models, specific resources can target at-risk populations with a moderate assurance of how many people to reach and where these resources should go.*

We thank the reviewer for their kind comments.

> *My little concern is about using different data from UNAIDS Key Population Atlas apart from the DHS, which is the may source data for the analyses. I believe the two variants of data are based on different designs, and combining them may not result in dependable results. It would have been more attainable if the UNAIDS data had been used in their sensitivity analysis to confirm the results from the DHS data.*

For the FSW risk group, we used age-disaggregated (Supplementary Figure A.4) national-level estimates from @stevens2022estimating to inform the national risk group population size within each age group, and household survey data to inform subnational variation.
Estimates of hidden populations like FSW from household surveys have significant limitations due in part to stigma around disclosing membership, as well as potential for not being included in the sampling frame [@abdul2014estimating].
For this reason, we believe it is more appropriate to calibrate our estimates to @stevens2022estimating, who as well as including the KP Atlas data, integrate data from other FSW population size estimates using a Bayesian mixed effects model.

Overall, we agree with the reviewer's comment that relying on data from different designs is not the ideal analysis, but we believe that it is the best approach using available data given known limitations identifying women who sell sex in household sampling frames and questionnaires.
We have noted this in the limitations section of the discussion.

> *Figure 1 is not clear. I recommend that the authors use a table as an alternative visualization.*

Table B.3 in the appendix provides an alternative tabulation of the surveys we used, with sample size further broken down by age group.
We are unsure on how to make Figure 1 might clearer.
Other examples of similar figures in published papers include @giguere2021trends (Figure 1), ...

\newpage

# Editorial comments {-}

> *Please send a completed Competing Interests' statement, including any COIs declared by your co-authors. If you have no competing interests to declare, please state: "The authors have declared that no competing interests exist". Otherwise please declare all competing interests beginning with the statement ``I have read the journal's policy and the authors of this manuscript have the following competing interests:*

Response here.

> *Please amend your detailed Financial Disclosure statement. This is published with the article. It must therefore be completed in full sentences and contain the exact wording you wish to be published. State what role the funders took in the study. If the funders had no role in your study, please state: "The funders had no role in study design, data collection and analysis, decision to publish, or preparation of the manuscript." If any authors received a salary from any of your funders, please state which authors and which funders. If you did not receive any funding for this study, please simply state: "The authors received no specific funding for this work."*

Response here.

> *Please ensure that the funders and grant numbers match between the Financial Disclosure field and the Funding Information tab in your submission form. Note that the funders must be provided in the same order in both places as well.*

Response here.

> *Please provide separate figure files in .tif or .eps format only and remove any figures embedded in your manuscript file. Please also ensure that all files are under our size limit of 10MB. For more information about figure files please see our guidelines: https://journals.plos.org/globalpublichealth/s/figures, https://journals.plos.org/globalpublichealth/s/figures#loc-file-requirements*

Response here.

> *Figure 2: please (a) provide a direct link to the base layer of the map (i.e., the country or region border shape) and ensure this is also included in the figure legend; and (b) provide a link to the terms of use / license information for the base layer image or shapefile. We cannot publish proprietary or copyrighted maps (e.g. Google Maps, Mapquest) and the terms of use for your map base layer must be compatible with our CC-BY 4.0 license. Note: if you created the map in a software program like R or ArcGIS, please locate and indicate the source of the basemap shapefile onto which data has been plotted. If your map was obtained from a copyrighted source please amend the figure so that the base map used is from an openly available source. Alternatively, please provide explicit written permission from the copyright holder granting you the right to publish the material under our CC-BY 4.0 license. Please note that the following CC BY licenses are compatible with PLOS license: CC BY 4.0, CC BY 2.0 and CC BY 3.0, meanwhile such licenses as CC BY-ND 3.0 and others are not compatible due to additional restrictions. If you are unsure whether you can use a map or not, please do reach out and we will be able to help you. The following websites are good examples of where you can source open access or public domain maps: U.S. Geological Survey (USGS) - All maps are in the public domain. (http://www.usgs.gov), PlaniGlobe - All maps are published under a Creative Commons license so please cite “PlaniGlobe, http://www.planiglobe.com, CC BY 2.0” in the image credit after the caption. (http://www.planiglobe.com/?lang=enl), Natural Earth - All maps are public domain. (http://www.naturalearthdata.com/about/terms-of-use/)*

Response here.

\newpage

# References {#references .unnumbered}