### 2023-04-08 src/docs_xx-05-23-hi-med/xx-05-23-hi-med.Rmd
Template for Hi-Med talk

---
title: Using statistics to respond to the HIV epidemic
subtitle: High Impact Medicine
author: Adam Howes
institute: Imperial College London
date: May 2023
bibliography: citations.bib
output:
  beamer_presentation:
    latex_engine: pdflatex
    highlight: haddock
    fig_width: 7 
    fig_height: 3
    includes:
        in_header: preamble.tex
---

## About me

* Maths undergraduate $\to$ statistics masters $\to$ statistics PhD
* Interested in impactful applications in global health and biosecurity
* Limited knowledge about epidemiology or biology

##

<!-- ## Background -->

<!-- ::: columns -->

<!-- :::: column -->

<!-- * In sub-Saharan Africa, adolescent girls and young women (AGYW) aged 15-29 are disproportionately at risk of HIV infection -->
<!--   * 28% of population and 44% of new infections -->
<!-- * This disparity is because of: -->
<!--   1. Younger age at first sex -->
<!--   2. Age patterns of sexual mixing -->
<!--   3. Structural vulnerabilities and power imbalances -->
<!--   4. Increased susceptibility to HIV infection -->

<!-- :::: -->

<!-- :::: column -->

<!-- ```{r echo=FALSE, out.width = "125px", fig.cap="Tweet from UNAIDS"} -->
<!-- knitr::include_graphics("tweet.png") -->
<!-- ``` -->

<!-- :::: -->

<!-- ::: -->

<!-- ## Prevention packages -->

<!-- * Prevention options can be divided into two: -->
<!--   1. Core package -->
<!--   2. Intensified interventions -->
<!-- * Insufficient resources to offer the more costly intensified interventions to all AGYW -->

<!-- \begin{center} -->
<!-- \begin{tcolorbox}[width=0.9\textwidth, colframe={title}, colback={white}, title={}] -->
<!-- $\implies$ How should interventions be prioritised to have the greatest impact? -->
<!-- \end{tcolorbox} -->
<!-- \end{center} -->

<!-- ## Stratified prevention -->

<!-- ::: columns -->

<!-- :::: column -->

<!-- * Global AIDS strategy 2021-2026 proposed stratifying HIV prevention for AGYW based on -->
<!--   1. Population-level HIV incidence -->
<!--   2. Individual-level sexual risk behaviour -->
<!-- * The two most proximal drivers of sexual transmission -->

<!-- :::: -->

<!-- :::: column -->

<!-- ```{r echo=FALSE, out.width = "125px", fig.cap="Global AIDS strategy"} -->
<!-- knitr::include_graphics("global-aids-strategy.png") -->
<!-- ``` -->

<!-- :::: -->

<!-- ::: -->

<!-- ## Scope for our work -->

<!-- ### Goals -->

<!-- 1. Enable implementation of prevention stratified by incidence and behaviour -->
<!-- 2. Assess the benefits of such approaches: is it worth it? -->

<!-- ### Approach -->

<!-- 1. Estimate the proportion of AGYW in four behavioural risk groups at a district level (in 13 countries identified as priority by The Global Fund) -->
<!-- 2. Analyze the new infections which could be reached by different stratified prevention strategies -->

<!-- ## Data -->

<!-- * Sexual behaviour data from household surveys to place respondents into $K = 4$ risk groups: -->
<!--   1. $k = 1$ Not sexually active -->
<!--   2. $k = 2$ One cohabiting sexual partner -->
<!--   3. $k = 3$ Non-regular sexual partner(s) -->
<!--   4. $k = 4$ Female sex workers -->
<!-- * District-level HIV incidence, prevalence, population size estimates from the Naomi model [@eaton2021naomi] -->
<!--   * Combines survey and programmatic data to estimate indicators in the general population -->
<!-- * Risk ratios from ALPHA network analysis [@slaymaker2020croi] and UNAIDS analysis led by Keith Sabin -->
<!--   * 10 longitudinal studies: https://alpha.lshtm.ac.uk/ -->

<!-- ## -->

<!-- ```{r echo=FALSE, out.width = "400px"} -->
<!-- knitr::include_graphics("depends/category-flowchart.pdf") -->
<!-- ``` -->

<!-- ## -->

<!-- ```{r echo=FALSE, out.width = "400px"} -->
<!-- knitr::include_graphics("depends/available-surveys.png") -->
<!-- ``` -->

<!-- ## Two-stage model for risk group proportions -->

<!-- \begin{center} -->
<!-- \begin{tcolorbox}[width=0.9\textwidth, colframe={title}, colback={white}, title={}] -->
<!-- How do we deal with the fact that only some of the surveys included a transactional sex question? This question is required to differentiate between the $k = 3$ and $k = 4$ risk groups -->
<!-- \end{tcolorbox} -->
<!-- \end{center} -->

<!-- * Our approach was to fit a two-stage model -->
<!--   1. Spatio-temporal multinomial logistic regression model for the proportion of AGYW in the $k = 1, 2, 3^{+}$ risk groups, using all 47 surveys -->
<!--   2. Spatial logistic regression model for the proportion of those in the $k = 3^{+} = \{3, 4\}$ risk groups who are in the $k = 4$ risk group, using only the 13 surveys with a specific transactional sex question -->

<!-- ## Notation -->

<!-- * $k \in \{1, \ldots, 4\}$: risk groups -->
<!-- * $i \in \{1, \ldots, n\}$: districts -->
<!-- * $c[i] \in \{\text{Botswana}, \ldots, \text{Zimbabwe}\}$: country containing district $i$ -->
<!-- * $t \in \{1999, \ldots, 2018\}$: years -->
<!-- * $a \in \{\text{15-19}, \text{20-24}, \text{25-29}\}$: age groups -->
<!-- * $y_{ita}^\star = (y_{ita1}^\star, y_{ita2}^\star, y_{ita3}^\star, y_{ita4}^\star)$: survey weighted multinomial observations -->
<!-- * $m_{ita}^\star$: survey weighted multinomial sample size -->

<!-- ## Multinomial logistic regression model -->

<!-- * Would like to use integrated nested Laplace approximations for fast, accurate inference, but `R-INLA` is not compatible with multinomial likelihoods because they depend on multiple elements of the latent field -->
<!-- * Instead, use that multinomial logistic regression models can be recast as a Poisson log-linear models using the Poisson trick -->
<!-- * This works because conditional on their sum, Poisson counts are jointly multinomially distributed -->
<!-- * If you include observation-specific random effects $\theta_{ita} \sim \mathcal{N}(0, 1000^2)$ in the linear predictor $\eta_{itak} = \theta_{ita} + \cdots$ then the sample sizes $m_{ita}^\star$ can be exactly recovered, ensuring the models are actually the same -->

<!-- ## Multinomial logistic regression model -->

<!-- \begin{align*} -->
<!--     y_{itak}^\star &\sim \text{xPoisson}(\lambda_{itak}) \\ -->
<!--     \log(\lambda_{itak}) &= \theta_{ita} + \beta_k + \zeta_{c[i]k} + \alpha_{ac[i]k} + \phi_{ik} + \gamma_{tk} + \delta_{itk}. -->
<!-- \end{align*} -->

<!-- * The terms are -->
<!--   * $\theta_{ita}$: observation (IID) -->
<!--   * $\beta_k$: category (IID) -->
<!--   * $\zeta_{ck}$: country-category (IID x IID) -->
<!--   * $\alpha_{ack}$: age-country-category (IID x IID x IID) -->
<!--   * $\phi_{ik}$: space-category ({IID, **Besag**} x IID) -->
<!--   * $\gamma_{tk}$: year-category ({**IID**, AR1} x IID) -->
<!--   * $\delta_{itk}$: space-year-category (Implemented but crashing on cluster at the moment...) -->
<!-- * Without category interactions, any effect will be washed out when we apply softmax to get the probabilities! -->

<!-- ##  -->

<!-- ```{r echo=FALSE, warning =FALSE, out.width="350px", fig.cap = "The Besag model, $\\phi_i \\, | \\, \\phi_{-i} \\sim \\mathcal{N} \\left(\\frac{1}{n_{\\delta i}} \\sum_{j: j \\sim i} \\phi_j, \\frac{1}{n_{\\delta i}\\tau_\\phi} \\right)$, operates on the neighbourhood graph."} -->
<!-- areas <- sf::read_sf("depends/zwe_areas.geojson") -->

<!-- areas %>% -->
<!--   filter(area_level == 2) %>% -->
<!--   bsae::plot_graph(add_geography = TRUE) -->
<!-- ``` -->

<!-- ## Multinomial logistic regression model -->

<!-- * Independent penalised complexity [@simpson2017penalising] priors on all standard deviation parameters with $\sigma = 0$ and $\mathbb{P}(\sigma > 2.5 = 0.01)$ -->
<!--   * Sidenote, I'm interested as to if joint priors might be more suitable -->
<!-- * Possible (but tricky) to define all these interactions in `R-INLA` by combination of the `group` and `replicate` options -->
<!--   * Check out the `aaa_fit_multi-sexbehav-sae` report if you're interested! -->
<!-- * Used sum-to-zero constraints to make posterior inferences interpretable -->
<!--   * Because we're interested in the contribution of each random effect to total variance -->
<!-- * Model comparison via CPO statistic -->

<!-- ## Logistic regression model -->

<!-- \begin{align*} -->
<!--     y_{ia4}^\star &\sim \text{Binomial} \left( y_{ia3}^\star + y_{ia4}^\star, q_{ia} \right), \label{eq:logistic-regression} \\ -->
<!--     q_{ia} &= \text{logit}^{-1} \left( \eta_{ia} \right), \\ -->
<!--     \eta_{ia} &= \beta_0 +  \zeta_{c[i]} + \alpha_{ac[i]} + \phi_i + \beta_{\texttt{cfsw}} \texttt{cfsw}_{c[i]}. -->
<!-- \end{align*} -->

<!-- * The terms are -->
<!--   * $\beta_0$: intercept -->
<!--   * $\zeta_{c[i]}$: country effects (IID) -->
<!--   * $\alpha_{ac[i]}$ age-country effects (IID) -->
<!--   * $\phi_i$: spatial effects (IID, **Besag**) -->
<!--   * Clients of FSW covariates (`cfswever`, **`cfswrecent`**) [@hodgins2022population] -->

<!-- ## Combination and FSW adjustment -->

<!-- * Take 1000 samples from each model, then manipulate suitably to generate estimates for all four risk groups -->
<!-- * We adjusted the samples from the $k = 4$ risk group to match age-country FSW estimates, reallocating into non-regular partner(s) -->
<!--   * Obtained these by disaggregating @stevens2022estimating by age using estimates of sexually active population from @nguyen2022trends -->

<!-- \begin{center} -->
<!-- \begin{tcolorbox}[width=0.9\textwidth, colframe={title}, colback={white}, title={}] -->
<!-- $\implies$ Estimates of risk group proportions $p_{itak}$ by district, year and age group -->
<!-- \end{tcolorbox} -->
<!-- \end{center} -->

<!-- ## -->

<!-- ```{r echo=FALSE, out.width="350px", fig.cap = "We found a geographic discontinuity in behaviour between Southern and Eastern Africa."} -->
<!-- knitr::include_graphics("depends/aids-abstract-A.pdf") -->
<!-- ``` -->

<!-- ## -->

<!-- ```{r echo=FALSE, out.width="350px", fig.cap = "Viewing the discontinuity another way."} -->
<!-- knitr::include_graphics("depends/aids-abstract-B.pdf") -->
<!-- ``` -->

<!-- ## -->

<!-- ```{r echo=FALSE, out.width="350px", fig.cap = "Proportions of variance explained."} -->
<!-- knitr::include_graphics("depends/aaa-variance-proportions.pdf") -->
<!-- ``` -->

<!-- ## Benefits of our modelled risk group estimates -->

<!-- * Integration of all relevant surveys -->
<!--   * Two-stage approach allowed estimating FSW proportion even for surveys without a specific transactional sex question -->
<!-- * Alleivating small-sample sizes by borrowing information -->
<!--   * We borrowed information across space, between countries and over surveys so that our estimates more plausibly reflect spatial heterogeneity -->
<!-- * Estimates where there isn't direct data -->
<!--   * Although some people think of this as "making up data", the data almost never "speak for themselves" (everything is a model) -->
<!--   * Uncertainty should be higher in regions with infilling: important to transparently communicate this -->

<!-- ## -->

<!-- ```{r echo=FALSE, out.width = "350px", fig.cap = "Illustration of the problem with direct survey estimates."} -->
<!-- knitr::include_graphics("depends/model-direct-benefits.png") -->
<!-- ``` -->

<!-- ## HIV prevalence and incidence by risk group -->

<!-- * We used our risk group proportion estimates together with incidence relative risk ratios and prevalence ratios to disaggregate general population HIV estimates in the most recent year -->
<!-- * Disaggregated number of new infections on a linear scale, and people living with HIV (PLHIV) on a logit scale -->
<!--   * Using a linear scale for PLHIV resulted in prevalences outside $[0, 1]$ -->

<!-- \begin{center} -->
<!-- \begin{tcolorbox}[width=0.9\textwidth, colframe={title}, colback={white}, title={}] -->
<!-- $\implies$ Estimates of HIV incidence $\lambda_{iak}$, number of new HIV infections $I_{iak}$, HIV prevalence $\rho_{iak}$ and PLHIV $H_{iak}$ by district, age group and risk group -->
<!-- \end{tcolorbox} -->
<!-- \end{center} -->

<!-- ## Prioritisation with risk group information -->

<!-- * Suppose we have all of the information (district, age, and risk group) -->
<!-- * Which are the strata with highest incidence? -->

<!-- ```{r include=FALSE} -->
<!-- df <- readr::read_csv("depends/incidence-district-sexbehav.csv") -->

<!-- df_area_age_behav <- df %>% -->
<!--   select(area_id, age_group, population_nosex12m, incidence_nosex12m, starts_with("population_sex"), starts_with("incidence_sex")) %>% -->
<!--   pivot_longer( -->
<!--     cols = starts_with("population") | starts_with("incidence"), -->
<!--     names_to = c("indicator", "category"), -->
<!--     names_sep = "_" -->
<!--   ) %>% -->
<!--   pivot_wider( -->
<!--     id_cols = c("area_id", "age_group", "category"), -->
<!--     names_from = "indicator", -->
<!--     values_from = "value" -->
<!--   ) -->
<!-- ``` -->

<!-- ```{r echo=FALSE} -->
<!-- df_area_age_behav %>% -->
<!--   arrange(desc(incidence)) %>% -->
<!--   head() %>% -->
<!--   knitr::kable(digits = 2) -->
<!-- ``` -->

<!-- ## Prioritisation without risk group information -->

<!-- * What about if we lost the risk group information? Now what are the strata with the highest incidence? -->

<!-- ```{r include=FALSE} -->
<!-- df_area_age <- df_area_age_behav %>% -->
<!--   group_by(area_id, age_group) %>% -->
<!--   summarise( -->
<!--     incidence = sum(incidence * population, na.rm = TRUE) / sum(population, na.rm = TRUE), -->
<!--     population = sum(population, na.rm = TRUE) -->
<!--   ) %>% -->
<!--   ungroup() -->
<!-- ``` -->

<!-- ```{r echo=FALSE} -->
<!-- df_area_age %>% -->
<!--   select(area_id, age_group, population, incidence) %>% -->
<!--   arrange(desc(incidence)) %>% -->
<!--   head() %>% -->
<!--   knitr::kable(digits = 2) -->
<!-- ``` -->

<!-- ## -->

<!-- ```{r echo=FALSE, out.width = "350px", fig.cap = "New infections reached prioritising according to different stratifications."} -->
<!-- knitr::include_graphics("depends/infections-reached.pdf") -->
<!-- ``` -->

<!-- ## Limitations -->

<!-- * Simplistic infections reached analysis -->
<!--   * No accounting for difficulty in reaching each strata -->
<!--   * Variable intervention effectiveness -->
<!--   * Change in strata membership -->
<!-- * Under-reporting of high risk sexual behaviours -->
<!--   * Variation in under-reporting (likely by age, foremost, and location, less so) particularly concerning -->
<!-- * Risk groups definition justification not clear -->
<!--   * Didn't consider other important characteristics that may determine risk e.g. condom usage -->
<!-- * Only focused on AGYW 15-29 -->
<!--   * Could be extended to adults of both sexes aged 15-49 -->

<!-- ## Takeaways -->

<!-- 1. Risk group estimates can help implement the Global AIDS Strategy; tool and user guide currently being rolled out! -->
<!-- 2. Importance of reaching FSW -->
<!-- 3. Countries have different epidemic profiles -->

<!-- ## Acknowledgements -->

<!-- * Joint work with members of the HIV inference group (`hiv-inference.org`) particularly Katie Risher and Jeff Eaton -->
<!-- * The code for this project is at `github.com/athowes/multi-agyw` -->

<!-- ## About me -->

<!-- * At Waterloo this term to work with Alex Stringer on improving the approximate inference method we use for the Naomi model -->
<!-- * Interested in Bayesian statistics, and applications in global health and biosecurity -->
<!-- * Recently worked on early detection of biological threats with metagenomic sequencing at the Nucleic Acid Observatory -->
<!-- * You can find me online at `athowes.github.io` -->
<!-- * Come say hi if you have any similar interests! -->

<!-- # Supplement -->

<!-- ## -->

<!-- ```{r echo=FALSE, out.width="350px", fig.cap = "Results of FSW age disaggregation."} -->
<!-- knitr::include_graphics("depends/age-disagg-fsw-line.pdf") -->
<!-- ``` -->

<!-- ##  -->

<!-- ```{r echo=FALSE, out.width="350px", fig.cap = "PIT histograms and ECDF difference plots."} -->
<!-- knitr::include_graphics("depends/coverage.pdf") -->
<!-- ``` -->

## Bibliography {.allowframebreaks}