### 2023-04-17 src/docs_17-05-23-hi-med/17-05-23-hi-med.Rmd
Start sketching out slides for #144

---
title: Using statistics to guide response to the HIV epidemic
subtitle: High Impact Medicine
author: Adam Howes
institute: Imperial College London
date: May 2023
bibliography: citations.bib
output:
  beamer_presentation:
    latex_engine: pdflatex
    highlight: haddock
    fig_width: 7 
    fig_height: 3
    includes:
        in_header: preamble.tex
---

## Outline

1. Background on small-area estimation of HIV indicators
2. Estimating HIV risk group proportions
3. Is this something you might want to work on?

# 1. Background on small-area estimation of HIV indicators

# 2. Estimating HIV risk group proportions

## Background

::: columns

:::: column

* In sub-Saharan Africa, adolescent girls and young women (AGYW) aged 15-29 are 28% of the population but 44% of new infections
* This disparity is because of:
  1. Younger age at first sex
  2. Age patterns of sexual mixing
  3. Structural vulnerabilities and power imbalances
  4. Increased susceptibility to HIV infection

::::

:::: column

```{r echo=FALSE, out.width = "125px", fig.cap="Tweet from UNAIDS"}
knitr::include_graphics("tweet.png")
```

::::

:::

## Prevention packages

* There are some prevention options ("core package") cheap enough they should be provided to all AGYW 
* Other options ("intensified interventions") are more costly, and there are insufficient resources to offer to all AGYW

\begin{center}
\begin{tcolorbox}[width=0.9\textwidth, colframe={title}, colback={white}, title={}]
$\implies$ How to prioritise interventions to have the greatest impact?
\end{tcolorbox}
\end{center}

## Stratified prevention

::: columns

:::: column

* Global AIDS strategy 2021-2026 proposed stratifying HIV prevention for AGYW based on
  1. Population-level HIV incidence
  2. Individual-level sexual risk behaviour
* These are the two most proximal drivers of sexual transmission

::::

:::: column

```{r echo=FALSE, out.width = "125px", fig.cap="Global AIDS strategy"}
knitr::include_graphics("global-aids-strategy.png")
```

::::

:::

# 3. Is this something you might want to work on?

## About me

* Maths undergraduate $\to$ statistics masters $\to$ statistics PhD
* Interested in impactful applications of statistics
* Limited domain knowledge in any given area! Need to collaborate!

## Many skills are required

* Mathematics and statistics
* Computer science and software engineering
* Epidemiology and public health
* Health economics and behavioural science
* Implementation science
* ...

## Thanks for listening!

* More information about me on my website `athowes.github.io`
* Feel free to get in contact if I can be useful at `ath19@ic.ac.uk`
* My PhD is supervised by Seth Flaxman and Jeff Eaton, and the risk group work is in collaboration with Imperial's HIV Inference Group and UNAIDS

```{r echo=FALSE, out.width = "300px", fig.align='center'}
knitr::include_graphics("mrc-gida-icl-mlgh.png")
```

## Bibliography {.allowframebreaks}