### 2023-05-15 src/docs_17-05-23-hi-med/17-05-23-hi-med.Rmd
Full outline draft of HiMed slides #144

---
title: Using statistics to guide response to the HIV epidemic
subtitle: High Impact Medicine
author: Adam Howes
institute: Imperial College London
date: May 2023
bibliography: citations.bib
output:
  beamer_presentation:
    latex_engine: pdflatex
    highlight: haddock
    fig_width: 7 
    fig_height: 3
    includes:
        in_header: preamble.tex
---

## Outline

1. Precision public health for HIV
2. Estimating risk group proportions
3. Might you want to do similar work?

# 1. Precision public health for HIV

## What is precision public health

* Precision medicine: the right intervention, at the right time, to the right **individual**
* Precision public health: the right intervention, at the right time, to the right **population**

## How can statistics help?

* Requires granular estimates of relevant quantities
* Disaggregated by age group, location, behaviour

## What should we estimate

* HIV prevalence: proportion of people who are infected
* HIV incidence: rate of new infections
* Antiretroviral therapy (ART) coverage: proportion of people living with HIV who take ART

##

```{r echo=FALSE, out.width = "300px", fig.cap="Population health goals."}
knitr::include_graphics("fast-track.png")
```

# 2. Estimating risk group proportions

##

```{r echo=FALSE}
cbpalette <- c("#56B4E9", "#009E73", "#E69F00", "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#999999")

data.frame(
  x = c("Population", "New infections"),
  y = c(0.28, 0.44)
) %>%
  ggplot(aes(x = x, y = y, fill = x)) +
    geom_col() +
    coord_flip() +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_manual(values = cbpalette) +
    labs(x = "", y = "", title = bquote("AGYW*" ~ bold("over-represented") ~ "in numbers of new infections"), caption = "*Adolescent girls and young women, 15-29") +
    guides(fill = "none") +
    theme_minimal()
```

## Why?

1. Younger age at first sex
2. Age patterns of sexual mixing
3. Structural vulnerabilities and power imbalances
4. Increased susceptibility to HIV infection

##

```{r echo=FALSE, out.width = "325px", fig.cap="HIV is a gender equality issue."}
knitr::include_graphics("rec5.png")
```

##

```{r echo=FALSE}
data.frame(
  price = 10,
  effectiveness = 40,
  name = "Condoms"
) %>%
  ggplot(aes(x = price, y = effectiveness, label = name)) +
    geom_text() +
    labs(x = "Price", y = "Effectiveness", title = "Not all interventions can* be provided to everyone", caption = "*Or should") +
    scale_x_continuous(breaks = c(10, 90), labels = c("Cheap", "Expensive"), limits = c(0, 100)) +
    scale_y_continuous(breaks = c(10, 90), labels = c("Works poorly", "Works well"), limits = c(0, 100)) +
    theme_minimal() +
    theme(
      panel.border = element_blank(),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.line = element_line(colour = "grey92", linewidth = 0.5, arrow = arrow(length = unit(5, "pt"), type = "open"))
    )
```

## How to prioritise interventions?

The **most proximal** drivers of transmission are
$$
\mathbb{P}(\text{transmission}) \propto \text{sexual partnerships} \times \mathbb{P}(\text{each partner can transmit})
$$

##

```{r echo=FALSE}
x <- y <- c(1:4)

expand.grid(x, y) %>%
  mutate(risk = ifelse(Var2 > 1, Var1 * Var2, 0)) %>%
  ggplot(aes(x = Var1, y = Var2)) +
  geom_tile(aes(fill = risk)) +
  scale_fill_gradientn(
    colours = c("white", "#E69F00"),
    breaks = c(2, 14),
    labels = c("Less risk", "More risk") 
  ) +
  scale_x_continuous(breaks = c(1:4), labels = c("<0.3%", "0.3-1%", "1-3%", ">3%")) +
  scale_y_continuous(breaks = c(1:4), labels = c("None", "Low", "High", "Very\n high")) +
  labs(x = "Population-level HIV incidence", y = "Individual-level risk behaviour", title = bquote("Important we consider" ~ bold("both") ~ "population setting and individual behaviour"), fill = "") +
  theme_minimal() +
  theme(
    panel.border = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(colour = "grey92", linewidth = 0.5, arrow = arrow(length = unit(5, "pt"), type = "open"))
  )
```

## Population-level HIV incidence

| Level       | District-level HIV incidence rate  |
|-------------|------------------------------------|
| Low         | <0.3%                              |
| Moderate    | 0.3% - 1.0%                        |
| High        | 1.0% - 3.0%                        |
| Very High   | >3.0%                              |

## Individual-level behavioural risk

| Level       | Behavioural risk group             | Risk ratio^[From...] |
|-------------|------------------------------------|------------|
| None        | Not sexually active                | 0          |
| Low         | One cohabiting partner             | 1          |
| High        | Multiple or non-regular partner(s) | 1.72       |
| Very High   | Female sex workers                 | 13         |

##

* Goal: enable prioritisation of HIV programming for AGYW
* Audience: programme managers, service providers and policy makers 
* Plan: estimate district-age behavioural risk group proportions, prevalences and incidences

##

* Describe data
* Describe model
* Spatio-temporal smoothing

##

```{r echo=FALSE, out.width = "350px", fig.cap = "Direct survey estimates are too noisy to work with, but our modelled estimates don't have the same issue."}
knitr::include_graphics("depends/model-direct-benefits.png")
```

##

* Results: geographic discontinuitiy

##

* Results: impact of prioritisation

##

* Where might this work go
* Yearly generation of estimates by country teams
* Extending to adults 15-49

##

```{r echo=FALSE, out.width = "300px", fig.align='center', fig.cap = "Example of user interface for estimates generation."}
knitr::include_graphics("naomi_user.png")
```

# 3. Might you want to do similar work?

##

Possible influence on health outcomes for many people **but** a part of a complex theory of change

\begin{align*}
\mathbb{P}(\text{impact}) = &\mathbb{P}(\text{relevant people interact with your work}) \\
&\times \mathbb{P}(\text{what you've done is useful}) \\
&\times \mathbb{P}(\text{message correctly conveyed}) \times \cdots \\
&\times \mathbb{P}(\text{intervention technically works}) \\
&\times \mathbb{P}(\text{intervention correctly implemented}) \times \cdots
\end{align*}

... hopefully you get the point

##

* Maths undergraduate $\to$ statistics masters $\to$ statistics PhD
* Involved with Effective Altruism during my masters, and since then I've been interested in impactful applications of statistics

##

* I have very limited domain knowledge! Important to collaborate
* Broad range of required skills: epidemiology, public health, health economics, behavioural science, implementation science, mathematics, statistics, computer science, software engineering

##

* How might the skills of medics fit in here?
* Need to think about this!

## Thanks for listening!

* More information on my website: `athowes.github.io/about`
* Feel free to get in contact if I can be useful: `ath19@ic.ac.uk`
* My PhD is supervised by Seth Flaxman and Jeff Eaton, and the risk group work is in collaboration with Imperial's HIV Inference Group, within the MRC Centre for Global Infectious Disease Analysis, and UNAIDS

```{r echo=FALSE, out.width = "300px", fig.align='center'}
knitr::include_graphics("mrc-gida-icl-mlgh.png")
```

##

```{r echo=FALSE, out.width = "300px", fig.align='center', fig.cap = "For more (informal!) take a look at the blog post on my website`"}
knitr::include_graphics("blog-post-updated.png")
```

##

```{r echo=FALSE, out.width = "300px", fig.align='center', fig.cap = "For more (formal!) take a look at the paper."}
knitr::include_graphics("plos-gph.png")
```

## Bibliography {.allowframebreaks}